### 2024-12-20 scratch.Rmd

---
title: "R Notebook"
output: html_notebook
---

@Turchin2001 argues convincingly for exponential (or geometric, if discrete) growth being the first law of population dyanmics.
It certainly is universal, at least as it is expressed in a similar fashion to Newton's laws of motion.
This paper also makes the case for why the logistic growth model is not so universal, and really just a special case.
But the basic $N = B - D$ is clear, and there is a nice counterfactual for spontaneous generation and how its existence would change this fundamental equation.
Basically, in our own equation below, we are making the assumption (without stating it, though we should) that individuals do not spontaneously generate; they must be introduced from existing individuals $i$ through birth.
@Turchin2001 does a good job of explaining why $N = B - D$ is really the best we do.
He distinguishes between postulates (axioms) and predictions (theorems), and he insists that "we need a set of general principles to place populations on a firm logical foundation".
I believe that we need a set of general principles to place population ecology and evolution on *the same* shared foundation.