### 2025-06-09 foundations_of_ecoevo.tex
Some more minor updates right before making final edits for Victor

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
]{article}
\usepackage{amsmath,amssymb}
\usepackage{setspace}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math} % this also loads fontspec
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else
  % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage[margin=1in]{geometry}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
% definitions for citeproc citations
\NewDocumentCommand\citeproctext{}{}
\NewDocumentCommand\citeproc{mm}{%
  \begingroup\def\citeproctext{#2}\cite{#1}\endgroup}
\makeatletter
 % allow citations to break across lines
 \let\@cite@ofmt\@firstofone
 % avoid brackets around text for \cite:
 \def\@biblabel#1{}
 \def\@cite#1#2{{#1\if@tempswa , #2\fi}}
\makeatother
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newenvironment{CSLReferences}[2] % #1 hanging-indent, #2 entry-spacing
 {\begin{list}{}{%
  \setlength{\itemindent}{0pt}
  \setlength{\leftmargin}{0pt}
  \setlength{\parsep}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
   \setlength{\leftmargin}{\cslhangindent}
   \setlength{\itemindent}{-1\cslhangindent}
  \fi
  % set entry spacing
  \setlength{\itemsep}{#2\baselineskip}}}
 {\end{list}}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{\hfill\break\parbox[t]{\linewidth}{\strut\ignorespaces#1\strut}}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
\usepackage{fancyhdr}
\usepackage{lineno}
\linenumbers
\modulolinenumbers[2]
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage{bookmark}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
  pdftitle={Foundations of ecological and evolutionary change},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\title{Foundations of ecological and evolutionary change}
\author{A. Bradley Duthie\(^{1,a,*}\) and Victor J. Luque\(^{2,a}\)}
\date{{[}1{]} Department of Biological and Environmental Sciences,
University of Stirling, Stirling, Scotland {[}2{]} Department of
Philosophy, University of Valencia, Valencia, Spain {[}*{]}
Corresponding author:
\href{mailto:alexander.duthie@stir.ac.uk}{\nolinkurl{alexander.duthie@stir.ac.uk}}
{[}a{]} Equal contribution}

\begin{document}
\maketitle

\setstretch{2}
\section{Abstract}\label{abstract}

Biological evolution is realised through the same mechanisms of birth
and death that underlie change in population density. The deep
interdependence between ecology and evolution is well-established, but
much theory in each discipline has been developed in isolation. While
recent work has accomplished eco-evolutionary synthesis, a gap remains
between the logical foundations of ecology and evolution. We bridge this
gap with a new equation that defines a summed value for a characteristic
across individuals in a population, from which the fundamental equations
of population ecology and evolutionary biology (the Price equation) are
derived. We thereby unify the fundamental equations of population
ecology and biological evolution under a general framework. Our
unification further demonstrates the equivalence between mean population
growth rate and evolutionary fitness and links this change to ecosystem
function. Finally, we outline how our proposed framework can be used to
unify social evolution and density-dependent population growth.

\textbf{Key words:} Ecology, Evolution, Eco-Evolutionary Theory,
Fundamental Theorem, Price Equation, Population Growth

\section{Introduction}\label{introduction}

Theoretical unification of phenomena is a powerful tool for scientific
advancement. Such unification has been a major goal in scientific
research throughout history (\citeproc{ref-Kitcher1993}{Kitcher 1993}),
and its value is perhaps most evident in reconciling unconnected models
and revealing new and unexpected empirical predictions. In evolutionary
biology, the Price equation (Box 1) provides a unifying framework for
evolutionary theory by exhaustively and exactly describing evolutionary
change for any closed evolving population
(\citeproc{ref-Price1970}{Price 1970}; \citeproc{ref-Luque2016}{Luque
2017}; \citeproc{ref-Lehtonen2020}{Lehtonen et al. 2020}). The Price
equation is therefore fundamental, in the sense that it binds together
all of evolutionary theory by formally defining what evolutionary change
is and is not (\citeproc{ref-Price1970}{Price 1970};
\citeproc{ref-Rice2004}{Rice 2004}; \citeproc{ref-Gardner2008}{Gardner
2008}; \citeproc{ref-Frank2015a}{Frank 2017};
\citeproc{ref-Luque2016}{Luque 2017}; \citeproc{ref-Luque2021}{Luque and
Baravalle 2021}). Using this formal definition, the scope of, and
relationships among, sub-disciplines within evolutionary theory can be
clarified. For example, fundamental equations of both population and
quantitative genetics can be derived from the Price equation
(\citeproc{ref-Queller2017}{Queller 2017}). This provides conceptual
clarity by demonstrating the logical consistency of different
theoretical frameworks within evolutionary biology. Our aim here is to
propose an equation that extends this conceptual clarity to include
population change and thereby provide a formal and exact definition of
joint ecological and evolutionary change.

In biological populations, ecological change is caused by the same
processes of individual birth and death that cause changes in allele
frequencies and phenotypes (\citeproc{ref-Turchin2001}{Turchin 2001};
\citeproc{ref-Connor2004}{Connor and Hartl 2004};
\citeproc{ref-Barfield2011}{Barfield et al. 2011}). As with evolution, a
fundamental equation can exhaustively and exactly define population
change. Unlike the Price equation, this fundamental equation is perhaps
self-evident; population change is simply the addition of births and
subtraction of deaths from current population size (\(N_{t}\)) to
recover the new population size (\(N_{t+1}\); Box 2). By definition,
this relationship of \(N_{t+1} = N_{t} + Births - Deaths\) applies to
any closed population. Turchin (\citeproc{ref-Turchin2001}{2001}) argues
that general principles are needed to establish a logical foundation for
population ecology, and this simple birth and death model and the
consequences that logically follow from it (e.g., exponential population
growth) is fundamental to population ecology. Any unifying definition of
joint ecological and evolutionary change must be able to derive both the
Price equation and this birth and death model.

The interdependence of ecological and evolutionary processes has long
been recognised (e.g., \citeproc{ref-Darwin1859}{Darwin 1859};
\citeproc{ref-Fisher1958}{Fisher 1958};
\citeproc{ref-Pelletier2009}{Pelletier et al. 2009}), but the rise of
eco-evolutionary models, which incorporate both, is relatively recent
following a widespread recognition that ecology and evolution can happen
on similar timescales (\citeproc{ref-Govaert2019}{Govaert et al. 2019};
\citeproc{ref-Yamamichi2023}{Yamamichi et al. 2023}). Currently, a
universally recognised formal definition of eco-evolutionary change is
lacking, with some theoreticians broadly interpreting ``eco-evolutionary
dynamics'' to allow for a separation of ecological and evolutionary
timescales (\citeproc{ref-Lion2023}{Lion et al. 2023}) and others
advocating for a more narrow interpretation in which no such separation
is permitted (\citeproc{ref-Bassar2021}{Bassar et al. 2021}). Bassar et
al. (\citeproc{ref-Bassar2021}{2021}) identify two types of
eco-evolutionary models that follow from these interpretations. The
first type uses separate equations to model population change versus
evolutionary change, thereby allowing for any number of ecological,
evolutionary, or environmental feedbacks (e.g.,
\citeproc{ref-Lion2018}{Lion 2018}; \citeproc{ref-Patel2018}{Patel et
al. 2018}; \citeproc{ref-Lion2023}{Lion et al. 2023}). The second type
models population demographics as functions of quantitative traits, with
ecological and evolutionary change following from demographic processes
and trait distributions operating on the same timescale (e.g.,
\citeproc{ref-Barfield2011}{Barfield et al. 2011};
\citeproc{ref-Simmonds2020}{Simmonds et al. 2020};
\citeproc{ref-Jaggi2024}{Jaggi et al. 2024}). Both model types can be
very general, but like all predictive models, they rely on simplifying
assumptions for tractability (\citeproc{ref-Levins1966}{Levins 1966};
\citeproc{ref-Luque2016}{Luque 2017}). These simplifying assumptions are
often grounded in the Price equation to demonstrate accuracy and logical
consistency when modelling evolutionary change (e.g.,
\citeproc{ref-Coulson2008}{Coulson and Tuljapurkar 2008};
\citeproc{ref-Barfield2011}{Barfield et al. 2011};
\citeproc{ref-Rees2016}{Rees and Ellner 2016};
\citeproc{ref-Lion2018}{Lion 2018}). For example, Barfield et al.
(\citeproc{ref-Barfield2011}{2011}) link their model back to Price
(\citeproc{ref-Price1970}{1970}), which they consider to be a
``universal law of evolution'', to place their conclusions concerning
stage structured evolution in the broader context of evolutionary
theory. The role of fundamental equations is therefore important for
unifying theory (\citeproc{ref-Luque2021}{Luque and Baravalle 2021}),
and we believe that a fundamental equation of eco-evolutionary change is
needed.

We present an equation from which the fundamental equations of ecology
and evolutionary biology can be derived. Derivation follows by adding
assumptions that are specific to population ecology or evolution in the
same way that key equations of population genetics (e.g., average
excess) or quantitative genetics (e.g., Breeder's equation) can be
derived from the Price equation (\citeproc{ref-Queller2017}{Queller
2017}). We propose our equation as a formal definition of
eco-evolutionary change.

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\begin{quote}
\textbf{Box 1:} The Price equation is an abstract formula to represent
evolutionary change. Formulated originally in the early 1970s by George
Price (\citeproc{ref-Price1970}{Price 1970},
\citeproc{ref-Price1972}{1972}), it postulates some basic properties
that all evolutionary systems must satisfy: change over time, ancestor
and descendant relations, and a character or phenotype
(\citeproc{ref-Rice2004}{Rice 2004}). Using simple algebraic language,
the Price equation represents evolutionary change with the predominant
notation,
\[\bar{w}\Delta\bar{z} = \mathrm{Cov}\left(w, z\right) + \mathrm{E}\left(w\Delta z\right).\]
In the above equation, \(\Delta\bar{z}\) is the change in the average
character value \(z\) over a time step of arbitrary length, \(w\) is an
individual's fitness, and \(\bar{w}\) average population fitness. On the
right-hand side of the equation, the first term is the covariance
between a character value \(z\) and fitness \(w\), which reflects
\(\bar{z}\) change attributable to differential survival and
reproduction. The second term is the expected value of \(\Delta z\),
which reflects the extent to which offspring deviate from parents in
\(z\) (\citeproc{ref-Rice2004}{Rice 2004};
\citeproc{ref-Okasha2006}{Okasha 2006}; \citeproc{ref-Frank2012}{Frank
2012}). A more specific version of the covariance term was already known
within the quantitative and population genetics tradition
(\citeproc{ref-Robertson1966}{Robertson 1966}), usually representing the
action of natural selection. The Price equation adds an expectation term
and abstracts away from any specific mechanisms of replication or
reproduction, or mechanisms of inheritance. Its definitional nature and
lack of substantive biological assumptions has been portrayed both as a
strength (\citeproc{ref-Rice2004}{Rice 2004};
\citeproc{ref-Frank2012}{Frank 2012}; \citeproc{ref-Luque2016}{Luque
2017}; \citeproc{ref-Baravalle2022}{Baravalle and Luque 2022}), and its
greatest weakness. The abstract nature of the Price equation places it
at the top of the hierarchy of fundamental theorems of evolution from
which the rest (Robertson's theorem, Fisher's fundamental theorem,
breeder's equation, Hamilton's rule, adaptive dynamics, etc.) can be
easily derived (\citeproc{ref-Lehtonen2016}{Lehtonen 2016},
\citeproc{ref-Lehtonen2018}{2018}; \citeproc{ref-Queller2017}{Queller
2017}). This abstractness is also key to developing a more general view
of evolution (\citeproc{ref-Rice2020}{Rice 2020};
\citeproc{ref-Luque2021}{Luque and Baravalle 2021};
\citeproc{ref-Edelaar2023}{Edelaar et al. 2023}). In contrast, some
researchers consider the Price equation just a triviality (even
tautological), and useless without further modelling assumptions
(\citeproc{ref-VanVeelen2005}{van Veelen 2005};
\citeproc{ref-VanVeelen2012}{van Veelen et al. 2012}). The debate
remains open (\citeproc{ref-VanVeelen2020}{van Veelen 2020};
\citeproc{ref-Baravalle2024}{Baravalle et al. 2025}).
\end{quote}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\begin{quote}
\textbf{Box 2:} The number of individuals in any closed population
(\(N\)) at any given time (\(t + 1\)) is determined by the existing
number of individuals (\(N_{t}\)), plus the number of births
(\(Births\)) minus the number of deaths (\(Deaths\)),
\[N_{t+1} = N_{t} + Births - Deaths.\] This equation is necessarily true
for any closed population. Despite its simplicity, it is a general
equation for defining population change and a starting point for
understanding population ecology. Turchin
(\citeproc{ref-Turchin2001}{2001}) notes that a consequence of this
fundamental equation is the tendency for populations to grow
exponentially (technically geometrically in the above case where time is
discrete). This inherent underlying tendency towards exponential growth
persists even as the complexities of real populations, such as
structure, stochasticity, or density-dependent effects are added to
population models (\citeproc{ref-Turchin2001}{Turchin 2001}). Given the
assumption that all individuals in the population are identical, a per
capita rate of birth, \(Births = bN_{t}\) and death, \(Deaths = dN_{t}\)
can be defined. Rearranging and defining \(\lambda = 1 + b - d\) gives,
\(N_{t+1} = N_{t}\lambda\). Here \(\lambda\) is the finite rate of
increase (\citeproc{ref-Gotelli2001}{Gotelli 2001}), and note that
because \(0 \leq d \leq 1\), \(\lambda \geq 0\). Verbally, the change in
size of any closed population equals its existing size times its finite
rate of increase.
\end{quote}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\section{A foundation for biological evolution and population
ecology}\label{a-foundation-for-biological-evolution-and-population-ecology}

To fully unify biological evolution and population ecology, we must
reconcile the the Price equation (Box 1) with the general equation for
population change (Box 2). The Price equation is critical for
partitioning different components of biological change
(\citeproc{ref-Price1970}{Price 1970}; \citeproc{ref-Frank1997}{Frank
1997}; \citeproc{ref-Gardner2008}{Gardner 2008};
\citeproc{ref-Luque2016}{Luque 2017}; \citeproc{ref-Queller2017}{Queller
2017}; \citeproc{ref-Lehtonen2018}{Lehtonen 2018}). It has also been
highly useful for integrating evolutionary theory across disciplines
(\citeproc{ref-Fox2006}{Fox 2006};
\citeproc{ref-Brantingham2007}{Brantingham 2007};
\citeproc{ref-MacCallum2012}{MacCallum et al. 2012};
\citeproc{ref-Frank2015}{Frank 2015}; \citeproc{ref-Godsoe2021}{Godsoe
et al. 2021}; \citeproc{ref-Ulrich2024}{Ulrich et al. 2024}). These
properties would seem to make it an intuitive starting point for a
logical foundation of ecology and evolution, perhaps through some kind
of mathematical equivalence (\citeproc{ref-Page2002}{Page and Nowak
2002}) or addition of terms (\citeproc{ref-Collins2009}{Collins and
Gardner 2009}), or through the use of its recursive structure
(\citeproc{ref-Kerr2009}{Kerr and Godfrey-Smith 2009};
\citeproc{ref-Frank2012}{Frank 2012}). But despite its flexibility, the
Price equation still relies on relative frequencies, which must by
definition sum to one (\citeproc{ref-Frank2015}{Frank 2015}). This is
because the Price equation describes the average change in a population;
the frequency of entities is scaled thereby conserving total probability
(\citeproc{ref-Frank2015}{Frank 2015}, \citeproc{ref-Frank2016}{2016}).
But to recover the fundamental principle of exponential population
growth (\citeproc{ref-Turchin2001}{Turchin 2001}), this scaling must be
avoided in a fundamental equation of ecology and evolution.

We therefore begin with the most fundamental axioms underlying the
ecology and evolution of living systems (\citeproc{ref-Rice2004}{Rice
2004}; \citeproc{ref-Rice2009}{Rice and Papadopoulos 2009}). In such
systems, diversity is discontinuous and can be defined in terms of
discrete entities (\citeproc{ref-Dobzhansky1970}{Dobzhansky 1970}). Our
framework is general enough that entities can be anything discrete, but
we will focus on each entity \(i\) as an individual organism.
Individuals give rise to new individuals through birth such that
\(\beta_{i}\) is the number of births attributable to \(i\). Individuals
are removed from the population through death such that \(\delta_{i}\)
is an indicator variable that takes a value of 1 (death of \(i\)) or 0
(persistence of \(i\)). All individuals are defined by some
characteristic \(z_{i}\), and \(\Delta z_{i}\) defines any change in
\(z_{i}\) from one time step \(t\) to the next \(t + 1\). The total
number of individuals in the population is \(N\). From this foundation,
we can define \(\Omega\) to be a summed characteristic across \(N\)
entities,

\[\Omega = \sum_{i=1}^{N} \left(\beta_{i} - \delta_{i} + 1 \right)\left(z_{i} + \Delta z_{i} \right).
\tag{1}
\]

The foundation of eco-evolutionary change is an interaction between the
demographic processes of birth and death
(\(\beta_{i} - \delta_{i} + 1\)) and some characteristic of an
individual (\(z_{i} + \Delta z_{i}\)). The value of \(\Omega\) is a
summation of individual characteristics, which takes the same units as
\(z\). From eqn 1, we can derive the most fundamental equations of
population ecology (Box 2) and evolutionary biology (Box 1) through an
appropriate interpretation of \(z\). Under some limited interpretations
of \(z\), we can also interpret \(\Omega\) as a metric of ecosystem
function.

\section{Population ecology}\label{population-ecology}

To recover the general equation for population ecology (Box 2), we
define \(z_{i}\) as the identity of \(i\) belonging to the population.
In other words, we set \(z_{i} = 1\) to simply indicate that \(i\) is a
member of the population. In this case, \(z\) is count, which takes the
unit 1 {[}note that `individual' is a label, not a unit; Newell and
Tiesinga (\citeproc{ref-NewellEtAl2019}{2019}){]}. Further, we assume
that individuals do not change species by setting \(\Delta z_{i} = 0\).
In this case,

\[\Omega = \sum_{i=1}^{N} \left(\beta_{i} - \delta_{i} + 1 \right).\]

We can now interpret \(\Omega\) as the population size at \(t+1\),
\(N_{t+1}\),

\[N_{t+1} = N_{t} + Births - Deaths.
\tag{2}
\]

If we assume that individuals are identical, then we can drop the
subscript \(i\) such that \(\beta_{i} = \beta\) and
\(\delta_{i} = \delta\) and note that \(N\beta\), \(N\delta\), and \(N\)
are the total births, total deaths, and size of the population at \(t\),
respectively,

Summing from 1 to \(N\), we can rewrite the above,

\[\Omega = N\left(1 + \beta - \delta \right).\]

If we define \(\lambda = 1 + \beta - \delta\) (Box 2), then we can
rewrite,

\[N_{t+1} = N_{t}\lambda.
\tag{3}
\]

We therefore recover the general equation for population ecology (eqn 2)
and the fundamental property of exponential growth in populations
(\citeproc{ref-Turchin2001}{Turchin 2001}) (eqn 3).

\section{Evolutionary biology}\label{evolutionary-biology}

Recovering the Price equation requires a few more steps. We start by
defining individual fitness,

\[w_{i} = \beta_{i} - \delta_{i} + 1.
\tag{4}
\]

In this definition, the longevity of the individual matters. An
individual that survives from \(t\) to \(t + 1\) has a higher fitness
than one that dies, even if both have the same reproductive output. With
this definition of fitness (eqn 4), we substitute,

\[\Omega = \sum_{i=1}^{N} \left(w_{i}z_{i} + w_{i}\Delta z_{i} \right).
\tag{5}
\]

We can break eqn 5 down further and multiply each side by \(1/N\),

\[\frac{1}{N}\Omega = \frac{1}{N}\sum_{i=1}^{N} \left(w_{i}z_{i} \right) + \frac{1}{N}\sum_{i=1}^{N}\left( w_{i}\Delta z_{i} \right).
\tag{6}
\]

We can rewrite the terms on the right-hand side of eqn 6 as expected
values and remove the subscripts,

\[\frac{1}{N}\Omega = \mathrm{E}\left(w z \right) + \mathrm{E}\left( w \Delta z  \right).
\tag{7}
\]

Now we must consider the total conservation of probability
(\citeproc{ref-Frank2015}{Frank 2015}, \citeproc{ref-Frank2016}{2016}).
In eqn 7, \(\Omega\) is the total sum trait values (\(z_{i}\)) across
the entire population at \(t + 1\) divided by the number of individuals
(\(N\)) in the population at \(t\). But the size of the population can
change from \(t\) to \(t + 1\). To recover mean trait change for the
Price equation (and therefore conserve total probability), we need to
account for this change in population size. The left-hand side of eqn 7
describes contributions to the sum trait value from the new population
at \(t + 1\). But we cannot treat \(\Omega/N\) as the mean of \(z\) at
\(t+1\) (\(\bar{z}'\)) because we need to weigh \(N\) by the mean
fitness of the population at \(t\) to account for any change in
population size from \(t\) to \(t+1\).

For example, if mean fitness at \(t\) is 2 (i.e., \(\bar{w} = 2\)), then
half as many individuals will have contributed to \(\Omega\) in
\(t + 1\) than would have if \(\bar{w} = 1\) (i.e., there are \(N\)
individuals at \(t\) and \(2N\) individuals at \(t + 1\)). We therefore
need to multiply the mean trait value \(\bar{z}'\) (at \(t + 1\)) by the
mean fitness \(\bar{w}\) (at \(t\)) to recover the mean contribution of
the \(N\) individuals at \(t\) to the total \(\Omega\) (this is noted in
the context of population genetics by \citeproc{ref-Ewens2014}{Ewens
2014}). Consequently,

\[\Omega = N\bar{w}\bar{z}'
\tag{8}
\]

Equation 8 conserves the total probability and now clarifies \(\Omega\)
as a summed trait value, which has the same units as \(z\) and equals
expected population growth at \(t\) times mean trait value at \(t + 1\).
This is consistent with the population ecology derivation from the
previous section where \(z_{i} = 1\) by definition, and
\(\Omega = N_{t+1}\) (note \(\lambda = \bar{w}\)). We can therefore
rewrite eqn 7,

\[\bar{w}\bar{z}' = \mathrm{E}\left(w z \right) + \mathrm{E}\left( w \Delta z  \right).
\tag{9}
\]

We can rearrange eqn 9 to derive the Price equation by expressing
covariance as,
\(\mathrm{Cov}(X,Y) = \mathrm{E}(XY) - \mathrm{E}(X)\mathrm{E}(Y)\), and
therefore
\(\mathrm{E}(XY) = \mathrm{Cov}(X,Y) + \mathrm{E}(X)\mathrm{E}(Y)\).
Substituting into eqn 9,

\[\bar{w}\bar{z}' = \mathrm{Cov}\left(w ,z \right) + \bar{w}\bar{z} + \mathrm{E}\left( w \Delta z  \right).\]

From here,

\[\bar{w}\left(\bar{z}' - \bar{z}\right) = \mathrm{Cov}\left(w ,z \right) + \mathrm{E}\left( w \Delta z  \right).\]

Since \(\Delta \bar{z} = \left(\bar{z}' - \bar{z}\right)\),

\[\bar{w}\Delta \bar{z} = \mathrm{Cov}\left(w ,z \right) + \mathrm{E}\left( w \Delta z  \right).
\tag{10}
\]

From eqn 1, which describes fundamental birth and death processes in a
population, we can derive both the most fundamental model of population
ecology (eqn 2; Box 2) and the fundamental equation of evolution (eqn
10; Box 1).

\section{Ecosystem function}\label{ecosystem-function}

In some cases, \(\Omega\) could also be intepreted as the total
contribution of a population to ecosystem function. This will be
restricted to cases in which \(z\) is an individual characteristic
defining an absolute quantity measured at the whole organism level, such
as biomass, seed production, carbon capture, flower visits, or nutrient
consumption. In these cases, the sum across individuals gives a
meaningful total quantity for the population. When \(z\) is instead
defined by relative organism-level measurements such wing loading,
nutrient ratio, or diet composition, or when \(z\) is measured at a
level of biological organisation below the organism (e.g., average cell
volume or leaf surface area), \(\Omega\) does not have a clear
population-level interpretation.

Box 3 provides a toy example of three plants with different fitnesses
and carbon captures. We use this toy example to demonstrate a concrete
example of our proposed framework.

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\begin{quote}
\textbf{Box 3:} As a toy example of our framework, consider a population
of \(N_{t} = 3\) annual plants in which individual fruit mass (kg) is
measured, and change is observed over a year. For all plants,
\(\delta_{i} = 1\), and let plant fecundities be \(\beta_{1} = 1\),
\(\beta_{2} = 1\), and \(\beta_{3} = 2\). Applying eqn 1 to population
change such that \(\Omega = N_{t+1}\), \(z_{i} = 1\) and
\(\Delta z_{i} = 0\),
\(N_{t+1} = (1 - 1 + 1)(1 + 0) + (1 - 1 + 1)(1 + 0) + (2 - 1 + 1)(1 + 0) = 4\)
(note \(\bar{w} = 4/3\), so \(N_{t}\bar{w} = 4\)). Focusing next on the
characteristic of total fruit mass, let \(z_{1} = 0.8\),
\(z_{2} = 1.0\), and \(z_{3} = 1.5\). Also let \(\Delta z_{i} = 0.1\)
for all plants to reflect a change in soil environment from \(t\) to
\(t + 1\). In this case, total fruit yield at \(t+1\) is
\(\Omega = (1 - 1 + 1)(0.8 + 0.1) + (1 - 1 + 1)(1.0 + 0.1) + (2 - 1 + 1)(1.5 + 0.1) = 5.2\)
kg. At \(t\), mean fruit yield per plant was 1.1 kg, but at \(t+1\),
mean fruit yield per plant is \(1.3\) kg. Note that
\(\mathrm{Cov}(w, z) = 2/15\) and \(\mathrm{E}(w \Delta z) = 2/15\), so
applying the Price equation, \(\bar{w}\Delta z = 2/15 + 2/15 = 4/15\).
Since \(\bar{w} = 4/3\), multiplying both sides of the equation by
\(3/4\) returns \(\Delta z = 0.2\), which is the mean difference in
fruit yield between \(t+1\) and \(t\). The framework expressed in eqn 1
thereby links population change, evolutionary change, and ecosystem
function.
\end{quote}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\section{Discussion}\label{discussion}

A classical sign of scientific progress is the ability to connect
disparate theories and models to show how empirical and theoretical
models are logical (mathematical) consequences of more fundamental ones
(\citeproc{ref-Nagel1961}{Nagel 1961};
\citeproc{ref-Morrison2000}{Morrison 2000}). Rather than making
simplifying assumptions, as is the approach for most ecological and
evolutionary models, we focus on fundamental axioms that are universal
to closed biological systems: discrete individuals, birth, death, and
change over time. We define an abstract sum (\(\Omega\)), to which all
individuals within the population contribute. From the simple
assumptions of population identity (\(z_{i} = 1\)) and invariability
(\(\Delta z_{i} = 0\)), we recover the most general equation of
population ecology (Box 2) and principle of exponential growth
(\(N_{t+1} = N_{t}\lambda\)). By defining individual fitness (\(w_{i}\))
and applying the total conservation of probability to individual
frequencies (\citeproc{ref-Frank2015}{Frank 2015},
\citeproc{ref-Frank2016}{2016}), we recover the most fundamental
equation of evolution (Box 1). Our eqn 1 thereby provides a foundation
for defining eco-evolutionary change in any population.

The Price equation provides a complete and exact description of
evolution in any closed evolving system (\citeproc{ref-Price1970}{Price
1970}; \citeproc{ref-Frank2012}{Frank 2012}). It is derived by
rearranging the mathematical notation defining changes in the
frequencies and characteristics of any type of entity
(\citeproc{ref-Price1970}{Price 1970};
\citeproc{ref-Gardner2008}{Gardner 2008}; \citeproc{ref-Luque2016}{Luque
2017}) (e.g., individuals, alleles). This derivation partitions total
characteristic change into different components, making it possible to
isolate evolutionary mechanisms (e.g., selection) and levels of
biological organisation (e.g., group, individual)
(\citeproc{ref-Frank1995}{Frank 1995}, \citeproc{ref-Frank2012}{2012};
\citeproc{ref-Kerr2009}{Kerr and Godfrey-Smith 2009};
\citeproc{ref-Luque2016}{Luque 2017}; \citeproc{ref-Okasha2020}{Okasha
and Otsuka 2020}). Because of its abstract nature and lack of any
system-specific assumptions, the Price equation is not dynamically
sufficient and makes no predictions about what will happen in any
particular system (\citeproc{ref-Gardner2020}{Gardner 2020}). Its role
is not to predict, but to formally and completely define and separate
components of evolutionary change. The same is true of the general
equation for population change (eqns 2 and 3), at least as we have used
it here where it serves to define what population growth means in
ecology. This equation formally and completely describes population
change in terms of births and deaths. In eqn 1, we therefore have a
fundamental equation from which we can derive complete ecological and
evolutionary change in any closed biological population, but one that is
revealed to be necessarily abstract and not dynamically sufficient. We
anticipate that this will be useful for eco-evolutionary theory in the
same way that the Price equation is useful for evolutionary theory:
facilitating specific model development and identifying new conceptual
insights, unresolved errors, and sources of model disagreements.

Our unification recovers the equivalence between the finite rate of
increase \(\lambda\) (Box 2) (\citeproc{ref-Gotelli2001}{Gotelli 2001})
and population mean evolutionary fitness \(\bar{w}\) (Box 1). The
population growth equation \(N_{t+1} = N_{t}\lambda\) can therefore be
rewritten as \(N_{t+1} = N_{t}\bar{w}\). This specific equivalence has
been proposed before (\citeproc{ref-Lande1976}{Lande 1976}), as has the
broader relationship between population growth rate and evolutionary
fitness (\citeproc{ref-Fisher1930}{Fisher 1930};
\citeproc{ref-Charlesworth1980}{Charlesworth 1980};
\citeproc{ref-Lande1982}{Lande 1982}; \citeproc{ref-Roff2008}{Roff
2008}; \citeproc{ref-Lion2018}{Lion 2018}). We show this from first
principles and clarify the relationship between fitness and population
growth. Over an arbitrary length of time, fitness is properly defined as
\(w_{i} = \beta_{i} - \delta_{i} + 1\). Over an individual's lifetime
(which, by definition, includes death), fitness is therefore
\(\beta_{i}\). The rate of change in ecology and evolution are reflected
in the first and second statistical moments of fitness, respectively.
Population growth rate reflects mean fitness \(\bar{w}\), while the rate
of evolutionary change reflects the variance in fitness
\(\mathrm{Var}(w)/\bar{w}\) (i.e., Fisher's fundamental theorem)
(\citeproc{ref-Frank1997}{Frank 1997}; \citeproc{ref-Rice2004}{Rice
2004}; \citeproc{ref-Queller2017}{Queller 2017}).

Our unification may also help explain, at least partially, some of the
success of classical population genetic models. For decades, population
genetics (and to some extent quantitative genetics) has been accused of
being a reductionist view of evolution, reducing everything to changes
in allele frequencies and abstracting away from individuals and their
environments (the ecological interactions)
(\citeproc{ref-MacColl2011}{MacColl 2011}). This has been a line of
argumentation by some defenders of the so-called Extended Synthesis
(\citeproc{ref-Pigliucci2009}{Pigliucci 2009}), especially in relation
to niche construction (\citeproc{ref-Odling-smee2003}{Odling-Smee et al.
2003}). Famously, Mayr (\citeproc{ref-Mayr1960}{1959}) characterized
population genetics as a simple input and output of genes, analogous to
``the adding of certain beans to a beanbag and the withdrawing of
others'' (also called ``beanbag genetics''). Historical critics of
population genetics could not articulate a clear explanation for why it
works so well despite all of its idealisations and simplifications. From
the Price equation, we are able to recover classical population and
quantitative genetic models (\citeproc{ref-Queller2017}{Queller 2017})
and develop new ones (\citeproc{ref-Rice2004}{Rice 2004},
\citeproc{ref-Rice2020}{2020}; \citeproc{ref-Luque2016}{Luque 2017};
\citeproc{ref-Lion2018}{Lion 2018}). Our eqn 1 contains ecology at its
core, and we show how the Price equation logically follows from it after
accounting for absolute population growth (eqn 8). We therefore conclude
that population and quantitative genetic equations contain ecology (no
matter how hidden), and the ecological nature of evolution is implicit
in population and quantitative genetic models.

We have focused on the dynamics of a closed population, and in doing so
leave ecological and evolutionary change attributable to migration for
future work. In population ecology, immigration and emigration can be
incorporated by adding a term for each to the right-hand side of the
equation in Box 2 (\citeproc{ref-Gotelli2001}{Gotelli 2001}). In
evolution, because the Price equation relies on mapping
ancestor-descendant relationships, accounting for migration is more
challenging. Kerr and Godfrey-Smith (\citeproc{ref-Kerr2009}{2009})
demonstrate how the Price equation can be extended to allow for
arbitrary links between ancestors and descendants, thereby extending the
Price equation to allow for immigration and emigration. Frank
(\citeproc{ref-Frank2012}{2012}) presents a simplified version of Kerr
and Godfrey-Smith (\citeproc{ref-Kerr2009}{2009}) that allows some
fraction of descendants to be unconnected to ancestors. In both ecology
and evolution, accounting for migration is done through the use of
additional terms on the right-hand side of the equations.

Our fundamental equation is complete and exact. It therefore implicitly
includes any effects of density dependence on population growth (see Box
2), or any social effects on evolutionary change (see Box 1). Both of
these effects can be made explicit by specifying how other individuals
in a population affect birth and death of a focal individual. Assuming
that effects of individuals are additive, independent, and identical, we
can derive well-established logistic growth equations and multi-level
selection from eqn 1 (Supporting Information S1).

We have shown that we can derive the fundamental equations of population
ecology and biological evolution from a single unifying equation.
Lastly, we propose our eqn 1 as a potential starting point for defining
ecosystem function and further conceptual unification between ecology,
evolution, and ecosystem function. The Price equation has previously
been used to investigate ecosystem function
(\citeproc{ref-Loreau2001}{Loreau and Hector 2001};
\citeproc{ref-Fox2006}{Fox 2006}), but not with any attempt towards
conceptual unification with evolutionary biology. For example, Fox
(\citeproc{ref-Fox2006}{2006}) applied the abstract properties of the
Price equation to partition total change in ecosystem function into
separate components attributable to species richness, species
composition, and context dependent effects. This approach provides a
framework for comparing the effects of biodiversity on ecosystem
function in empirical systems (\citeproc{ref-Fox2006}{Fox 2006};
\citeproc{ref-Winfree2015}{Winfree et al. 2015};
\citeproc{ref-Mateo-Tomas2017}{Mateo-Tomás et al. 2017}). Instead, our
eqn 1 defines \(\Omega\) as total ecosystem function contributed by a
focal population. It is therefore possible to investigate ecological,
evolutionary, and ecosystem function change from the same shared
framework.

\section{Acknowledgements}\label{acknowledgements}

This manuscript was supported by joint funding between the French
Foundation for Research on Biodiversity (FRB) Centre for the Synthesis
and Analysis of Biodiversity (CESAB) and the German Centre for
Integrative Biodiversity Research (sDiv). It was written as part of the
Unification of Modern Coexistence Theory and Price Equation (UNICOP)
project. Victor J. Luque was also supported by the Spanish Ministry of
Science and Innovation (Project: PID2021-128835NB-I00), and the
Conselleria d'Innovaci\(\'{o}\), Universitats, Ci\(\`{e}\)ncia i
Societat Digital -- Generalitat Valenciana (Project: CIGE/2023/16). We
are grateful for many conversations with S\(\'{e}\)bastien Lion, Kelsey
Lyberger, Swati Patel, and especially Lynn Govaert, whose questions
helped us clarify the relationship between population growth and
fitness. Brad Duthie would also like to thank Brent Danielson and Stan
Harpole. Victor J. Luque would also like to thank Lorenzo Baravalle, Pau
Carazo, Santiago Ginnobili, Manuel Serra, and Ariel Roff\(\'{e}\).

\section{Author Contributions}\label{author-contributions}

Both authors came up with the question idea. Following many discussions
between the authors, ABD proposed the initial equation with subsequent
exploration and development from both authors. Both authors contributed
to the writing.

\section{Competing Interests}\label{competing-interests}

The authors declare no competing interests.

\section{Data Availability}\label{data-availability}

This work does not include any data.

\section*{References}\label{references}
\addcontentsline{toc}{section}{References}

\phantomsection\label{refs}
\begin{CSLReferences}{0}{0}
\bibitem[\citeproctext]{ref-Baravalle2022}
Baravalle, L., and V. J. Luque. 2022.
\href{https://doi.org/10.1387/theoria.21940}{Towards a pricean
foundation for cultural evolutionary theory}. Theoria 37:209--231.

\bibitem[\citeproctext]{ref-Baravalle2024}
Baravalle, L., A. Roffé, V. J. Luque, and S. Ginnobili. 2025.
\href{https://doi.org/10.1007/s13752-024-00482-4}{The value of price}.
Pages 12--24 \emph{in}Biological Theory (Vol. 20).

\bibitem[\citeproctext]{ref-Barfield2011}
Barfield, M., R. D. Holt, and R. Gomulkiewicz. 2011.
\href{https://doi.org/10.1086/658903}{Evolution in stage-structured
populations}. American Naturalist 177:397--409.

\bibitem[\citeproctext]{ref-Bassar2021}
Bassar, R. D., T. Coulson, J. Travis, and D. N. Reznick. 2021.
\href{https://doi.org/10.1111/ele.13712}{Towards a more precise -- and
accurate -- view of eco-evolution}. Ecology Letters 24:623--625.

\bibitem[\citeproctext]{ref-Brantingham2007}
Brantingham, P. J. 2007. \href{https://doi.org/10.2307/40035853}{{A
unified evolutionary model of archaeological style and function based on
the Price equation}}. American Antiquity 72:395--416.

\bibitem[\citeproctext]{ref-Charlesworth1980}
Charlesworth, B. 1980. Evolution in age-structured populations.
Cambridge studies in mathematical biology. Cambridge University Press,
Cambridge.

\bibitem[\citeproctext]{ref-Collins2009}
Collins, S., and A. Gardner. 2009.
\href{https://doi.org/10.1111/j.1461-0248.2009.01340.x}{{Integrating
physiological, ecological and evolutionary change: A Price equation
approach}}. Ecology Letters 12:744--757.

\bibitem[\citeproctext]{ref-Connor2004}
Connor, J., and D. L. Hartl. 2004. {A premier of ecological genetics}.
Sinauer Associates Incorporated.

\bibitem[\citeproctext]{ref-Coulson2008}
Coulson, T., and S. Tuljapurkar. 2008.
\href{https://doi.org/10.1086/591693}{The dynamics of a quantitative
trait in an age-structured population living in a variable environment}.
American Naturalist 172:599--612.

\bibitem[\citeproctext]{ref-Darwin1859}
Darwin, C. 1859. The origin of species. Penguin.

\bibitem[\citeproctext]{ref-Dobzhansky1970}
Dobzhansky, T. 1970. Genetics of the evolutionary process (Vol. 139).
Columbia University Press.

\bibitem[\citeproctext]{ref-Edelaar2023}
Edelaar, Pim, J. Otsuka, and V. J. Luque. 2023.
\href{https://doi.org/10.1111/brv.12910}{{A generalised approach to the
study and understanding of adaptive evolution}}. Biological Reviews
98:352--375.

\bibitem[\citeproctext]{ref-Ewens2014}
Ewens, W. J. 2014.
\href{https://doi.org/10.1007/s10539-013-9412-0}{Grafen, the price
equations, fitness maximization, optimisation and the fundamental
theorem of natural selection}. Biology and Philosophy 29:197--205.

\bibitem[\citeproctext]{ref-Fisher1930}
Fisher, R. A. 1930. The genetical theory of natural selection. Oxford
University Press, Oxford, UK.

\bibitem[\citeproctext]{ref-Fisher1958}
Fisher, R. A. 1958. The genetical theory of natural selection (2nd ed.).
Dover.

\bibitem[\citeproctext]{ref-Fox2006}
Fox, J. W. 2006.
\href{https://doi.org/10.1890/0012-9658(2006)87\%5B2687:utpetp\%5D2.0.co;2}{{Using
the price equation to partition the effects of biodiversity loss on
ecosystem function}}. Ecology 87:2687--2696.

\bibitem[\citeproctext]{ref-Frank1995}
Frank, S. A. 1995. \href{https://doi.org/10.1006/jtbi.1995.0148}{{George
Price's contributions to evolutionary genetics}}. Journal of Theoretical
Biology 175:373--388.

\bibitem[\citeproctext]{ref-Frank1997}
---------. 1997.
\href{https://doi.org/10.1111/j.1558-5646.1997.tb05096.x}{{The Price
equation, Fisher's fundamental theorem, kin selection, and causal
analysis}}. Evolution 51:1712--1729.

\bibitem[\citeproctext]{ref-Frank2012}
---------. 2012.
\href{https://doi.org/10.1111/j.1420-9101.2012.02498.x}{{Natural
selection. IV. The Price equation}}. Journal of Evolutionary Biology
25:1002--1019.

\bibitem[\citeproctext]{ref-Frank2015}
---------. 2015. \href{https://doi.org/10.3390/e17107087}{{D'Alembert's
direct and inertial forces acting on populations: The price equation and
the fundamental theorem of natural selection}}. Entropy 17:7087--7100.

\bibitem[\citeproctext]{ref-Frank2016}
---------. 2016. \href{https://doi.org/10.3390/e18050192}{{Common
probability patterns arise from simple invariances}}. Entropy 18:1--22.

\bibitem[\citeproctext]{ref-Frank2015a}
---------. 2017. \href{https://doi.org/10.1002/ece3.2922}{{Universal
expressions of population change by the Price equation: natural
selection, information, and maximum entropy production}}. Ecology and
Evolution 1--16.

\bibitem[\citeproctext]{ref-Gardner2008}
Gardner, A. 2008. \href{https://doi.org/10.1016/j.cub.2008.01.005}{{The
Price equation}}. Current Biology 18:198--202.

\bibitem[\citeproctext]{ref-Gardner2020}
---------. 2020. \href{https://doi.org/10.1098/rstb.2019.0361}{{Price's
equation made clear}}. Philosophical Transactions of the Royal Society
B: Biological Sciences 375:20190361.

\bibitem[\citeproctext]{ref-Godsoe2021}
Godsoe, W., K. E. Eisen, D. Stanton, and K. M. Sirianni. 2021.
\href{https://doi.org/10.1007/s12080-020-00478-3}{Selection and
biodiversity change}. Theoretical Ecology 14:367--379.

\bibitem[\citeproctext]{ref-Gotelli2001}
Gotelli, N. J. 2001. A primer of ecology. Sinauer associate. Inc.
Sunderland, MA.

\bibitem[\citeproctext]{ref-Govaert2019}
Govaert, L., E. A. Fronhofer, S. Lion, C. Eizaguirre, D. Bonte, M. Egas,
A. P. Hendry, et al. 2019.
\href{https://doi.org/10.1111/1365-2435.13241}{{Eco-evolutionary
feedbacks---Theoretical models and perspectives}}. Functional Ecology
33:13--30.

\bibitem[\citeproctext]{ref-Jaggi2024}
Jaggi, H., W. Zuo, R. Kentie, J. M. Gaillard, T. Coulson, and S.
Tuljapurkar. 2024. \href{https://doi.org/10.1111/ele.14551}{Density
dependence shapes life-history trade-offs in a food-limited population}.
Ecology letters 27:e14551.

\bibitem[\citeproctext]{ref-Kerr2009}
Kerr, B., and P. Godfrey-Smith. 2009.
\href{https://doi.org/10.1111/j.1558-5646.2008.00570.x}{{Generalization
of the price equation for evolutionary change}}. Evolution 63:531--536.

\bibitem[\citeproctext]{ref-Kitcher1993}
Kitcher, P. 1993. {The advancement of science}. Oxford University Press,
New York.

\bibitem[\citeproctext]{ref-Lande1976}
Lande, R. 1976. Natural selection and random genetic drift in phenotypic
evolution. Evolution 30:314--334.

\bibitem[\citeproctext]{ref-Lande1982}
---------. 1982. \href{https://doi.org/10.2307/1936778}{{A quantitative
genetic theory of life history evolution}}. Ecology 63:607--615.

\bibitem[\citeproctext]{ref-Lehtonen2016}
Lehtonen, J. 2016.
\href{https://doi.org/10.1016/j.tree.2016.07.006}{{Multilevel selection
in kin selection language}}. Trends in Ecology and Evolution xx:1--11.

\bibitem[\citeproctext]{ref-Lehtonen2018}
---------. 2018. \href{https://doi.org/10.1086/694891}{{The Price
equation, gradient dynamics, and continuous trait game theory}}.
American Naturalist 191:146--153.

\bibitem[\citeproctext]{ref-Lehtonen2020}
Lehtonen, J., S. Okasha, and H. Helanterä. 2020.
\href{https://doi.org/10.1098/rstb.2019.0350}{{Fifty years of the Price
equation}}. Philosophical Transactions of the Royal Society B:
Biological Sciences 375:20190350.

\bibitem[\citeproctext]{ref-Levins1966}
Levins, R. 1966. \href{https://doi.org/10.2307/27836590}{The strategy of
model building in population biology}. American Naturalist.

\bibitem[\citeproctext]{ref-Lion2018}
Lion, S. 2018. \href{https://doi.org/10.1086/694865}{{Theoretical
approaches in evolutionary ecology: environmental feedback as a unifying
perspective}}. American Naturalist 191.

\bibitem[\citeproctext]{ref-Lion2023}
Lion, S., A. Sasaki, and M. Boots. 2023.
\href{https://doi.org/10.1111/ele.14183}{Extending eco-evolutionary
theory with oligomorphic dynamics}. Ecology Letters 26:S22--S46.

\bibitem[\citeproctext]{ref-Loreau2001}
Loreau, M., and A. Hector. 2001.
\href{https://doi.org/10.1038/35083573}{{Partitioning selection and
complementarity in biodiversity experiments}}. Nature 412:72--76.

\bibitem[\citeproctext]{ref-Luque2016}
Luque, V. J. 2017. \href{https://doi.org/10.1007/s10539-016-9538-y}{{One
equation to rule them all: a philosophical analysis of the Price
equation}}. Biology and Philosophy 32:1--29.

\bibitem[\citeproctext]{ref-Luque2021}
Luque, V. J., and L. Baravalle. 2021.
\href{https://doi.org/10.1007/s11229-021-03339-6}{{The mirror of
physics: on how the price equation can unify evolutionary biology}}.
Synthese 199:12439--12462.

\bibitem[\citeproctext]{ref-MacCallum2012}
MacCallum, R. M., M. Mauch, A. Burt, and A. M. Leroi. 2012.
\href{https://doi.org/10.5061/dryad.h0228}{{Evolution of music by public
choice}}. Proceedings of the National Academy of Sciences
109:12081--12086.

\bibitem[\citeproctext]{ref-MacColl2011}
MacColl, A. D. C. 2011.
\href{https://doi.org/10.1016/j.tree.2011.06.009}{{The ecological causes
of evolution}}. Trends in Ecology and Evolution 26:514--522.

\bibitem[\citeproctext]{ref-Mateo-Tomas2017}
Mateo-Tomás, P., P. P. Olea, M. Moleón, N. Selva, and J. A.
Sánchez-Zapata. 2017. \href{https://doi.org/10.1111/geb.12673}{{Both
rare and common species support ecosystem services in scavenger
communities}}. Global Ecology and Biogeography 26:1459--1470.

\bibitem[\citeproctext]{ref-Mayr1960}
Mayr, E. 1959. \href{https://doi.org/10.1101/SQB.1959.024.01.003}{Where
are we? Genetics and twentieth century darwinism}. Pages 1--14
\emph{in}Cold spring harbor symposia on quantitative biology (Vol. 24).

\bibitem[\citeproctext]{ref-Morrison2000}
Morrison, M. 2000. Unifying scientific theories: Physical concepts and
mathematical structures. Cambridge University Press.

\bibitem[\citeproctext]{ref-Nagel1961}
Nagel, E. 1961. The structure of science: Problems in the logic of
scientific explanation. Harcourt, Brace \& World, New York, NY, USA.

\bibitem[\citeproctext]{ref-NewellEtAl2019}
Newell, D. B., and E. Tiesinga. 2019.
\href{https://doi.org/10.6028/NIST.SP.330-2019}{The international system
of units (SI)}.

\bibitem[\citeproctext]{ref-Odling-smee2003}
Odling-Smee, F. J., K. N. Laland, and M. W. Feldman. 2003. Niche
construction: The neglected process in evolution. Princeton University
Press.

\bibitem[\citeproctext]{ref-Okasha2006}
Okasha, S. 2006.
\href{https://doi.org/10.1093/acprof:oso/9780199267972.001.0001}{Evolution
and the levels of selection}. Oxford University Press.

\bibitem[\citeproctext]{ref-Okasha2020}
Okasha, S., and J. Otsuka. 2020.
\href{https://doi.org/10.1098/rstb.2019.0365}{{The Price equation and
the causal analysis of evolutionary change}}. Philosophical Transactions
of the Royal Society B: Biological Sciences 375:20190365.

\bibitem[\citeproctext]{ref-Page2002}
Page, K. M., and M. A. Nowak. 2002.
\href{https://doi.org/10.1006/jtbi.2002.3112}{Unifying evolutionary
dynamics}. Journal of theoretical biology 219:93--98.

\bibitem[\citeproctext]{ref-Patel2018}
Patel, S., M. H. Cortez, and S. J. Schreiber. 2018.
\href{https://doi.org/10.1101/104505}{{Partitioning the effects of
eco-evolutionary feedbacks on community stability}}. American Naturalist
191:1--29.

\bibitem[\citeproctext]{ref-Pelletier2009}
Pelletier, F., D. Garant, and A. P. Hendry. 2009.
\href{https://doi.org/10.1098/rstb.2009.0027}{Eco-evolutionary
dynamics}. Philosophical Transactions of the Royal Society London B
364:1483--1489.

\bibitem[\citeproctext]{ref-Pigliucci2009}
Pigliucci, M. 2009.
\href{https://doi.org/10.1111/j.1749-6632.2009.04578.x}{{An extended
synthesis for evolutionary biology}}. Annals of the New York Academy of
Sciences 1168:218--228.

\bibitem[\citeproctext]{ref-Price1970}
Price, G. R. 1970. \href{https://doi.org/10.1038/227520a0}{{Selection
and covariance}}.

\bibitem[\citeproctext]{ref-Price1972}
---------. 1972.
\href{https://doi.org/10.1111/j.1469-1809.1957.tb01874.x}{{Extension of
covariance selection mathematics}}. Annals of Human Genetics
35:485--490.

\bibitem[\citeproctext]{ref-Queller2017}
Queller, D. C. 2017. \href{https://doi.org/10.1086/690937}{{Fundamental
theorems of evolution}}. American Naturalist 189:000--000.

\bibitem[\citeproctext]{ref-Rees2016}
Rees, M., and S. P. Ellner. 2016.
\href{https://doi.org/10.1111/2041-210X.12487}{Evolving integral
projection models: Evolutionary demography meets eco-evolutionary
dynamics}. Methods in Ecology and Evolution 7:157--170.

\bibitem[\citeproctext]{ref-Rice2004}
Rice, S. H. 2004. {Evolutionary theory: mathematical and conceptual
foundations}. Sinauer Associates Incorporated.

\bibitem[\citeproctext]{ref-Rice2020}
---------. 2020.
\href{https://doi.org/10.1098/rstb.2019.0353}{{Universal rules for the
interaction of selection and transmission in evolution}}. Philosophical
Transactions of the Royal Society B: Biological Sciences 375.

\bibitem[\citeproctext]{ref-Rice2009}
Rice, S. H., and A. Papadopoulos. 2009.
\href{https://doi.org/10.1371/journal.pone.0007130}{{Evolution with
stochastic fitness and stochastic migration}}. PLoS One 4.

\bibitem[\citeproctext]{ref-Robertson1966}
Robertson, A. 1966. \href{https://doi.org/10.1017/S0003356100037752}{{A
mathematical model of the culling process in dairy cattle}}. Animal
Science 8:95--108.

\bibitem[\citeproctext]{ref-Roff2008}
Roff, D. A. 2008.
\href{https://doi.org/10.1007/s12041-008-0056-9}{{Defining fitness in
evolutionary models}}. Journal of Genetics 87:339--348.

\bibitem[\citeproctext]{ref-Simmonds2020}
Simmonds, E. G., E. F. Cole, B. C. Sheldon, and T. Coulson. 2020.
\href{https://doi.org/10.1111/ele.13603}{Phenological asynchrony: A
ticking time-bomb for seemingly stable populations?} Ecology Letters.
Blackwell Publishing Ltd.

\bibitem[\citeproctext]{ref-Turchin2001}
Turchin, P. 2001.
\href{https://doi.org/10.1034/j.1600-0706.2001.11310.x}{Does population
ecology have general laws?} Oikos 94:17--26.

\bibitem[\citeproctext]{ref-Ulrich2024}
Ulrich, W., N. J. Gotelli, G. Strona, and W. Godsoe. 2024.
\href{https://doi.org/10.1016/j.ecolmodel.2024.110695}{Reconsidering the
price equation: Benchmarking the analytical power of additive
partitioning in ecology}. Ecological Modelling 491:110695.

\bibitem[\citeproctext]{ref-VanVeelen2005}
van Veelen, M. 2005.
\href{https://doi.org/10.1016/j.jtbi.2005.04.026}{{On the use of the
Price equation}}. Journal of Theoretical Biology 237:412--426.

\bibitem[\citeproctext]{ref-VanVeelen2020}
---------. 2020. \href{https://doi.org/10.1098/rstb.2019.0355}{{The
problem with the Price equation}}. Philosophical Transactions of the
Royal Society B: Biological Sciences 375:20190355.

\bibitem[\citeproctext]{ref-VanVeelen2012}
van Veelen, M., J. García, M. W. Sabelis, and M. Egas. 2012.
\href{https://doi.org/10.1016/j.jtbi.2011.07.025}{{Group selection and
inclusive fitness are not equivalent; the Price equation vs. models and
statistics}}. Journal of Theoretical Biology 299:64--80.

\bibitem[\citeproctext]{ref-Winfree2015}
Winfree, R., J. W. Fox, N. M. Williams, J. R. Reilly, and D. P.
Cariveau. 2015. \href{https://doi.org/10.1111/ele.12424}{{Abundance of
common species, not species richness, drives delivery of a real-world
ecosystem service}}. Ecology Letters 18:626--635.

\bibitem[\citeproctext]{ref-Yamamichi2023}
Yamamichi, M., S. P. Ellner, and N. G. Hairston. 2023.
\href{https://doi.org/10.1111/ele.14197}{Beyond simple adaptation:
Incorporating other evolutionary processes and concepts into
eco-evolutionary dynamics}. Ecology Letters 26:S16--S21.

\end{CSLReferences}

\end{document}