### 2025-06-18 foundations_of_ecoevo_arxiv.tex

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
]{article}
\usepackage{amsmath,amssymb}
\usepackage{setspace}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math} % this also loads fontspec
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else
  % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage[margin=1in]{geometry}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
% definitions for citeproc citations
\NewDocumentCommand\citeproctext{}{}
\NewDocumentCommand\citeproc{mm}{%
  \begingroup\def\citeproctext{#2}\cite{#1}\endgroup}
\makeatletter
 % allow citations to break across lines
 \let\@cite@ofmt\@firstofone
 % avoid brackets around text for \cite:
 \def\@biblabel#1{}
 \def\@cite#1#2{{#1\if@tempswa , #2\fi}}
\makeatother
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newenvironment{CSLReferences}[2] % #1 hanging-indent, #2 entry-spacing
 {\begin{list}{}{%
  \setlength{\itemindent}{0pt}
  \setlength{\leftmargin}{0pt}
  \setlength{\parsep}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
   \setlength{\leftmargin}{\cslhangindent}
   \setlength{\itemindent}{-1\cslhangindent}
  \fi
  % set entry spacing
  \setlength{\itemsep}{#2\baselineskip}}}
 {\end{list}}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{\hfill\break\parbox[t]{\linewidth}{\strut\ignorespaces#1\strut}}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
\usepackage{fancyhdr}
\usepackage{lineno}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage{bookmark}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
  pdftitle={Foundations of ecological and evolutionary change},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\title{Foundations of ecological and evolutionary change}
\author{A. Bradley Duthie\(^{1,a,*}\) and Victor J. Luque\(^{2,a}\)}
\date{{[}1{]} Department of Biological and Environmental Sciences,
University of Stirling, Stirling, Scotland {[}2{]} Department of
Philosophy, University of Valencia, Valencia, Spain {[}*{]}
Corresponding author:
\href{mailto:alexander.duthie@stir.ac.uk}{\nolinkurl{alexander.duthie@stir.ac.uk}}
{[}a{]} Equal contribution}

\begin{document}
\maketitle

\setstretch{1}
\section{Abstract}\label{abstract}

Biological evolution is realised through the same mechanisms of birth
and death that underlie change in population density. The deep
interdependence between ecology and evolution is well-established, and
recent models focus on integrating eco-evolutionary dynamics to
demonstrate how ecological and evolutionary processes interact and feed
back upon each other. Nevertheless, a gap remains between the logical
foundations of ecology and evolution. Both population ecology and
evolution have fundamental equations that define how the size of a
population (ecology) and the average characteristic within a population
(evolution) change, respectively, over time. These fundamental equations
are a complete and exact description of change for any closed
population, but how they are formally linked remains unclear. We link
the fundamental equations of population ecology and evolution with an
equation that sums how individual characteristics interact with
individual fitness in a population. From this equation, we derive the
fundamental equations of population ecology and evolutionary biology
(the Price equation). We thereby identify an overlooked bridge between
ecology and biological evolution. Our unification formally recovers the
equivalence between mean population growth rate and evolutionary fitness
and links this change to ecosystem function. We outline how our
framework can be used to further develop eco-evolutionary theory.

\textbf{Key words:} Ecology, Evolution, Eco-Evolutionary Theory,
Fundamental Theorem, Price Equation, Population Growth

\section{Introduction}\label{introduction}

Theoretical unification is a powerful tool for scientific advancement.
Such unification has been a major goal in scientific research throughout
history (\citeproc{ref-Smocovitis1992}{Smocovitis 1992};
\citeproc{ref-Kitcher1993}{Kitcher 1993}), and its value is perhaps most
evident in reconciling unconnected models and revealing new and
unexpected empirical predictions. In evolutionary biology, the Price
equation (Box 1) provides a unifying framework for evolutionary theory
by exhaustively and exactly describing evolutionary change for any
closed evolving population (\citeproc{ref-Price1970}{Price 1970};
\citeproc{ref-Luque2016}{Luque 2017};
\citeproc{ref-Lehtonen2020}{Lehtonen et al. 2020}). The Price equation
is therefore fundamental, in the sense that it binds together all of
evolutionary theory by formally defining what evolutionary change is and
is not (\citeproc{ref-Price1970}{Price 1970};
\citeproc{ref-Rice2004}{Rice 2004}; \citeproc{ref-Gardner2008}{Gardner
2008}; \citeproc{ref-Frank2015a}{Frank 2017};
\citeproc{ref-Luque2016}{Luque 2017}; \citeproc{ref-Luque2021}{Luque and
Baravalle 2021}). Using this formal definition, the scope of, and
relationships among, sub-disciplines within evolutionary theory can be
clarified. For example, fundamental equations of both population and
quantitative genetics can be derived from the Price equation
(\citeproc{ref-Queller2017}{Queller 2017}). This provides conceptual
clarity by demonstrating the logical consistency of different
theoretical frameworks within evolutionary biology. Our aim here is to
propose an equation that extends this conceptual clarity to include
population size change and thereby provide a formal and exact definition
for joint ecological and evolutionary change.

In biological populations, ecological change is caused by the same
processes of individual birth and death that cause evolutionary changes
in allele frequencies and phenotypes (\citeproc{ref-Turchin2001}{Turchin
2001}; \citeproc{ref-Connor2004}{Connor and Hartl 2004};
\citeproc{ref-Barfield2011}{Barfield et al. 2011}). As with evolution, a
fundamental equation can exhaustively and exactly define population
change. Unlike the Price equation, this fundamental equation is perhaps
self-evident. Population change is simply the addition of individuals
minus the removal of individuals from current population size
(\(N_{t}\)), which recovers the new population size (\(N_{t+1}\); Box
2). By definition, this relationship of
\(N_{t+1} = N_{t} + Births - Deaths\) applies to any closed population.
Turchin (\citeproc{ref-Turchin2001}{2001}) argues that general
principles are needed to establish a logical foundation for population
ecology, and this simple birth and death model and the consequences that
logically follow from it (e.g., exponential population growth) are
fundamental to population ecology. Any unifying definition of joint
ecological and evolutionary change must be able to derive both the Price
equation and this birth and death model.

The union of ecological and evolutionary processes has long been
recognised (e.g., \citeproc{ref-Darwin1859}{Darwin 1859};
\citeproc{ref-Fisher1958}{Fisher 1958};
\citeproc{ref-Pelletier2009}{Pelletier et al. 2009}), but the rise of
eco-evolutionary models, which incorporate both, is relatively recent
following a widespread recognition that ecology and evolution can happen
on similar timescales (\citeproc{ref-Govaert2019}{Govaert et al. 2019};
\citeproc{ref-Yamamichi2023}{Yamamichi et al. 2023}). Currently, a
universally recognised formal definition of eco-evolutionary change is
lacking, with some theoreticians broadly interpreting ``eco-evolutionary
dynamics'' to allow for a separation of ecological and evolutionary
timescales (\citeproc{ref-Lion2023}{Lion et al. 2023}) and others
advocating for a more narrow interpretation in which no such separation
is permitted (\citeproc{ref-Bassar2021}{Bassar et al. 2021}). Bassar et
al. (\citeproc{ref-Bassar2021}{2021}) identify two types of
eco-evolutionary models that follow from these interpretations. The
first type uses separate equations to model population size change
versus evolutionary change, thereby allowing for any number of
ecological, evolutionary, or environmental feedbacks (e.g.,
\citeproc{ref-Lion2018}{Lion 2018}; \citeproc{ref-Patel2018}{Patel et
al. 2018}; \citeproc{ref-Lion2023}{Lion et al. 2023}). The second type
models population demographics as functions of quantitative traits with
ecological and evolutionary change following from demographic processes
and trait distributions operating on the same timescale (e.g.,
\citeproc{ref-Barfield2011}{Barfield et al. 2011};
\citeproc{ref-Simmonds2020}{Simmonds et al. 2020};
\citeproc{ref-Jaggi2024}{Jaggi et al. 2024}). Both model types can be
very general, but like all predictive models, they rely on simplifying
assumptions for tractability (\citeproc{ref-Levins1966}{Levins 1966};
\citeproc{ref-Luque2016}{Luque 2017}). These simplifying assumptions are
often grounded in the Price equation to demonstrate accuracy and logical
consistency when modelling evolutionary change (e.g.,
\citeproc{ref-Coulson2008}{Coulson and Tuljapurkar 2008};
\citeproc{ref-Barfield2011}{Barfield et al. 2011};
\citeproc{ref-Rees2016}{Rees and Ellner 2016};
\citeproc{ref-Lion2018}{Lion 2018}). For example, Barfield et al.
(\citeproc{ref-Barfield2011}{2011}) link their model back to Price
(\citeproc{ref-Price1970}{1970}), which they consider to be a
``universal law of evolution'', to place their conclusions concerning
stage structured evolution in the broader context of evolutionary
theory. The role of fundamental equations is therefore important for
unifying theory (\citeproc{ref-Luque2021}{Luque and Baravalle 2021}),
and we believe that a fundamental equation of eco-evolutionary change
has been curiously overlooked.

We present an equation from which the fundamental equations of ecology
and evolutionary biology can be derived. Derivation follows by adding
assumptions that are specific to population ecology or evolution in the
same way that key equations of population genetics or quantitative
genetics can be derived from the Price equation by restricting the
domain of interest (e.g., to allele frequencies in the case of
population genetics, or to continuous phenotypes in the case of
quantitative genetics, \citeproc{ref-Queller2017}{Queller 2017}). We
propose our equation as a formal definition of eco-evolutionary change.

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\begin{quote}
\textbf{Box 1:} The Price equation is an abstract formula to represent
evolutionary change. Formulated originally in the early 1970s by George
Price (\citeproc{ref-Price1970}{Price 1970},
\citeproc{ref-Price1972}{1972}), it postulates some basic properties
that all evolutionary systems must satisfy: change over time, ancestor
and descendant relations, and a character or phenotype
(\citeproc{ref-Rice2004}{Rice 2004}). Using simple algebraic language,
the Price equation represents evolutionary change with the predominant
notation,
\[\bar{w}\Delta\bar{z} = \mathrm{Cov}\left(w, z\right) + \mathrm{E}\left(w\Delta z\right).\]
In the above equation, \(\Delta\bar{z}\) is the change in the average
character value \(z\) over a time step of arbitrary length, \(w\) is an
individual's fitness, and \(\bar{w}\) average population fitness. On the
right-hand side of the equation, the first term is the covariance
between a character value \(z\) and fitness \(w\), which reflects
\(\bar{z}\) change attributable to differential survival and
reproduction. The second term is the expected value of \(\Delta z\),
which reflects the extent to which offspring deviate from parents in
\(z\) (\citeproc{ref-Rice2004}{Rice 2004};
\citeproc{ref-Okasha2006}{Okasha 2006}; \citeproc{ref-Frank2012}{Frank
2012}). A more specific version of the covariance term was already known
within the quantitative and population genetics tradition
(\citeproc{ref-Robertson1966}{Robertson 1966}), usually representing the
action of natural selection. The Price equation adds an expectation term
and abstracts away from any specific mechanisms of replication or
reproduction, or mechanisms of inheritance. Its definitional nature and
lack of substantive biological assumptions has been portrayed both as a
strength (\citeproc{ref-Rice2004}{Rice 2004};
\citeproc{ref-Frank2012}{Frank 2012}; \citeproc{ref-Luque2016}{Luque
2017}; \citeproc{ref-Baravalle2022}{Baravalle and Luque 2022}), and its
greatest weakness. The abstract nature of the Price equation places it
at the top of the hierarchy of fundamental theorems of evolution from
which the rest (Robertson's theorem, Fisher's fundamental theorem,
breeder's equation, Hamilton's rule, adaptive dynamics, etc.) can be
derived (\citeproc{ref-Lehtonen2016}{Lehtonen 2016},
\citeproc{ref-Lehtonen2018}{2018}, \citeproc{ref-Lehtonen2020a}{2020};
\citeproc{ref-Queller2017}{Queller 2017}). This abstractness is also key
to developing a more general view of evolution
(\citeproc{ref-Rice2020}{Rice 2020}; \citeproc{ref-Luque2021}{Luque and
Baravalle 2021}; \citeproc{ref-Edelaar2023}{Edelaar et al. 2023}). In
contrast, some researchers consider the Price equation just a triviality
(even tautological), and useless without further modelling assumptions
(\citeproc{ref-VanVeelen2005}{van Veelen 2005};
\citeproc{ref-VanVeelen2012}{van Veelen et al. 2012}). The debate
remains open (\citeproc{ref-VanVeelen2020}{van Veelen 2020};
\citeproc{ref-Baravalle2024}{Baravalle et al. 2025}).
\end{quote}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\begin{quote}
\textbf{Box 2:} The number of individuals in any closed population
(\(N\)) at any given time (\(t + 1\)) is determined by the number of
individuals at \(t\) (\(N_{t}\)), plus the number of births (\(Births\))
minus the number of deaths (\(Deaths\)),
\[N_{t+1} = N_{t} + Births - Deaths.\] This equation is necessarily true
for any closed population. Despite its simplicity, it is a general
equation for defining population change and a starting point for
understanding population ecology. Turchin
(\citeproc{ref-Turchin2001}{2001}) notes that a consequence of this
fundamental equation is the tendency for populations to grow
exponentially (technically geometrically in the above case where time is
discrete). This inherent underlying tendency towards exponential growth
persists even as the complexities of real populations, such as
structure, stochasticity, or density-dependent effects are added to
population models (\citeproc{ref-Turchin2001}{Turchin 2001}). Given the
assumption that all individuals in the population are identical, a per
capita rate of birth, \(Births = bN_{t}\), and death,
\(Deaths = dN_{t}\), can be defined. Rearranging and defining
\(\lambda = 1 + b - d\) gives, \(N_{t+1} = N_{t}\lambda\). Here
\(\lambda\) is the finite rate of increase
(\citeproc{ref-Gotelli2001}{Gotelli 2001}), and note that because
\(0 \leq d \leq 1\), \(\lambda \geq 0\). Verbally, the change in size of
any closed population equals its existing size times its finite rate of
increase.
\end{quote}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\section{A foundation for biological evolution and population
ecology}\label{a-foundation-for-biological-evolution-and-population-ecology}

To unify biological evolution and population ecology, we must reconcile
the Price equation (Box 1) with the general equation for population
change (Box 2). The Price equation is critical for partitioning
different components of biological change
(\citeproc{ref-Price1970}{Price 1970}; \citeproc{ref-Frank1997}{Frank
1997}; \citeproc{ref-Gardner2008}{Gardner 2008};
\citeproc{ref-Luque2016}{Luque 2017}; \citeproc{ref-Queller2017}{Queller
2017}; \citeproc{ref-Lehtonen2018}{Lehtonen 2018},
\citeproc{ref-Lehtonen2020a}{2020}). It has also been highly useful for
integrating evolutionary theory across disciplines
(\citeproc{ref-Fox2006}{Fox 2006};
\citeproc{ref-Brantingham2007}{Brantingham 2007};
\citeproc{ref-MacCallum2012}{MacCallum et al. 2012};
\citeproc{ref-Frank2015}{Frank 2015}; \citeproc{ref-Godsoe2021}{Godsoe
et al. 2021}; \citeproc{ref-Ulrich2024}{Ulrich et al. 2024}). These
properties would seem to make it an intuitive starting point for a
logical foundation of ecology and evolution, perhaps through some kind
of mathematical equivalence (\citeproc{ref-Page2002}{Page and Nowak
2002}) or addition of terms (\citeproc{ref-Collins2009}{Collins and
Gardner 2009}), or through the use of its recursive structure
(\citeproc{ref-Kerr2009}{Kerr and Godfrey-Smith 2009};
\citeproc{ref-Frank2012}{Frank 2012}). But despite its flexibility, the
Price equation still relies on relative frequencies, which must by
definition sum to one (\citeproc{ref-Frank2015}{Frank 2015}). This is
because the Price equation describes the average change in a population;
the frequency of entities is scaled thereby conserving total probability
(\citeproc{ref-Frank2015}{Frank 2015}, \citeproc{ref-Frank2016}{2016}).
But to recover the fundamental principle of exponential population
growth (\citeproc{ref-Turchin2001}{Turchin 2001}), this scaling must be
avoided.

We therefore begin with the most fundamental axioms underlying the
ecology and evolution of living systems (\citeproc{ref-Rice2004}{Rice
2004}; \citeproc{ref-Rice2009}{Rice and Papadopoulos 2009}). In such
systems, diversity is discontinuous, in the sense that living systems
are composed of discrete entities including individual organisms and
groups of organisms (\citeproc{ref-Dobzhansky1970}{Dobzhansky 1970}).
Our framework is general enough that entities can be anything discrete,
but we will focus on each entity \(i\) as an individual organism.
Individuals give rise to new individuals through birth such that
\(\beta_{i}\) is the number of births attributable to \(i\). Individuals
are removed from the population through death such that \(\delta_{i}\)
is an indicator variable that takes a value of 1 (death of \(i\)) or 0
(persistence of \(i\)). All individuals are defined by some
characteristic \(z_{i}\), and \(\Delta z_{i}\) defines any change in
\(z_{i}\) from one time step \(t\) to the next \(t + 1\). The total
number of individuals in the population is \(N\). From this foundation,
we can define \(\Omega\) to be a summed characteristic across \(N\)
entities,

\[\Omega = \sum_{i=1}^{N} \left(\beta_{i} - \delta_{i} + 1 \right)\left(z_{i} + \Delta z_{i} \right).
\tag{1}
\]

This foundation of eco-evolutionary change is a statistical interaction
between the demographic processes of birth and death
(\(\beta_{i} - \delta_{i} + 1\)) and some characteristic of an
individual (\(z_{i} + \Delta z_{i}\)). The value of \(\Omega\) is a
summation of individual characteristics and takes the same units as
\(z\). From eqn 1, we can derive the most fundamental equations of
population ecology and evolutionary biology through an appropriate
interpretation of \(z\). Under more limited interpretations of \(z\), we
can also interpret \(\Omega\) as a metric of ecosystem function.

\section{Population ecology}\label{population-ecology}

To recover the general equation for population ecology (Box 2), we
define \(z_{i}\) as the identity of \(i\) belonging to the population.
In other words, we set \(z_{i} = 1\) to simply indicate that \(i\) is
one member of the population. In this restricted case, \(z\) is count,
which takes the unit 1 (note that `individual' is a label, not a unit,
see \citeproc{ref-NewellEtAl2019}{Newell and Tiesinga 2019}). Further,
we assume that individuals do not change species by setting
\(\Delta z_{i} = 0\). In this case,

\[\Omega = \sum_{i=1}^{N} \left(\beta_{i} - \delta_{i} + 1 \right).\]

We can now interpret \(\Omega\) as the population size at \(t+1\),
\(N_{t+1}\). Summing up \(\beta_{i}\), \(\delta_{i}\), and current
individuals tallies up the total number of individuals in the next time
step,

\[N_{t+1} = N_{t} + Births - Deaths.
\tag{2}
\]

If we assume that individuals are identical, then we can drop the
subscript \(i\) such that \(\beta_{i} = \beta\) and
\(\delta_{i} = \delta\) and note that \(N\beta\), \(N\delta\), and \(N\)
are the total births, total deaths, and size of the population at \(t\),
respectively,

Summing from 1 to \(N\), we can rewrite the above,

\[\Omega = N\left(1 + \beta - \delta \right).\]

If we define \(\lambda = 1 + \beta - \delta\) (Box 2), then we can
rewrite,

\[N_{t+1} = N_{t}\lambda.
\tag{3}
\]

We therefore recover the general equation for population ecology (eqn 2)
and the fundamental property of exponential growth in populations (eqn
3, \citeproc{ref-Turchin2001}{Turchin 2001}).

\section{Evolutionary biology}\label{evolutionary-biology}

Recovering the Price equation requires a few more steps. We start by
defining individual fitness,

\[w_{i} = \beta_{i} - \delta_{i} + 1.
\tag{4}
\]

In this definition, the longevity of the individual matters. By
definition, an individual that survives from \(t\) to \(t + 1\) has a
higher fitness than one that dies, even if both have the same
reproductive output. With this definition of fitness (eqn 4), we
substitute,

\[\Omega = \sum_{i=1}^{N} \left(w_{i}z_{i} + w_{i}\Delta z_{i} \right).
\tag{5}
\]

We can break eqn 5 down further and multiply each side by \(1/N\),

\[\frac{1}{N}\Omega = \frac{1}{N}\sum_{i=1}^{N} \left(w_{i}z_{i} \right) + \frac{1}{N}\sum_{i=1}^{N}\left( w_{i}\Delta z_{i} \right).
\tag{6}
\]

We can rewrite the terms on the right-hand side of eqn 6 as expected
values and remove the subscripts,

\[\frac{1}{N}\Omega = \mathrm{E}\left(w z \right) + \mathrm{E}\left( w \Delta z  \right).
\tag{7}
\]

Now we must consider the total conservation of probability
(\citeproc{ref-Frank2015}{Frank 2015}, \citeproc{ref-Frank2016}{2016}).
In eqn 7, \(\Omega\) is the total sum trait values (\(z_{i}\)) across
the entire population at \(t + 1\) divided by the number of individuals
(\(N\)) in the population at \(t\). But the size of the population can
change from \(t\) to \(t + 1\). To recover mean trait change for the
Price equation (and therefore conserve total probability), we need to
account for this change in population size. The left-hand side of eqn 7
describes contributions to the sum trait value from the new population
at \(t + 1\). But we cannot treat \(\Omega/N\) as the mean of \(z\) at
\(t+1\) (\(\bar{z}'\)) because we need to weight \(N\) by the mean
fitness of the population at \(t\) to account for any change in
population size from \(t\) to \(t+1\).

For example, if mean fitness at \(t\) is 2 (i.e., \(\bar{w} = 2\)), then
half as many individuals will have contributed to \(\Omega\) in
\(t + 1\) than would have if \(\bar{w} = 1\) (i.e., there are \(N\)
individuals at \(t\) and \(2N\) individuals at \(t + 1\)). We therefore
need to multiply the mean trait value \(\bar{z}'\) (at \(t + 1\)) by the
mean fitness \(\bar{w}\) (at \(t\)) to recover the mean contribution of
the \(N\) individuals at \(t\) to the total \(\Omega\) (this has been
applied in the context of population genetics by
\citeproc{ref-Ewens2014}{Ewens 2014}). Consequently,

\[\Omega = N\bar{w}\bar{z}'
\tag{8}
\]

Equation 8 conserves the total probability and now clarifies \(\Omega\)
as a summed trait value, which has the same units as \(z\) and equals
expected population growth at \(t\) times mean trait value at \(t + 1\).
This is consistent with the population ecology derivation from the
previous section where \(z_{i} = 1\) by definition, and
\(\Omega = N_{t+1}\) (note \(\lambda = \bar{w}\)). We can therefore
rewrite eqn 7,

\[\bar{w}\bar{z}' = \mathrm{E}\left(w z \right) + \mathrm{E}\left( w \Delta z  \right).
\tag{9}
\]

We can rearrange eqn 9 to derive the Price equation by expressing
covariance as
\(\mathrm{Cov}(X,Y) = \mathrm{E}(XY) - \mathrm{E}(X)\mathrm{E}(Y)\), and
therefore
\(\mathrm{E}(XY) = \mathrm{Cov}(X,Y) + \mathrm{E}(X)\mathrm{E}(Y)\).
Substituting into eqn 9,

\[\bar{w}\bar{z}' = \mathrm{Cov}\left(w ,z \right) + \bar{w}\bar{z} + \mathrm{E}\left( w \Delta z  \right).\]

From here,

\[\bar{w}\left(\bar{z}' - \bar{z}\right) = \mathrm{Cov}\left(w ,z \right) + \mathrm{E}\left( w \Delta z  \right).\]

Since \(\Delta \bar{z} = \left(\bar{z}' - \bar{z}\right)\),

\[\bar{w}\Delta \bar{z} = \mathrm{Cov}\left(w ,z \right) + \mathrm{E}\left( w \Delta z  \right).
\tag{10}
\]

Equation 10 is the Price equation. From eqn 1, which describes
fundamental birth and death processes in a population, we can therefore
derive both the most fundamental model of population ecology (eqn 2; Box
2) and the fundamental equation of evolution (eqn 10; Box 1).

\section{Ecosystem function}\label{ecosystem-function}

In some cases, \(\Omega\) could also be interpreted as the total
contribution of a population to ecosystem function. This is restricted
to cases in which \(z\) is a characteristic defining an absolute
quantity measured at the whole organism level such as biomass, seed
production, carbon capture, flower visits, or nutrient consumption. In
such cases, the sum across individuals gives a meaningful total quantity
for the population (e.g., the total biomass or seeds produced in the
population). When \(z\) is instead defined by relative organism-level
measurements such wing loading, nutrient ratio, or diet composition, or
when \(z\) is measured at a level of biological organisation below the
organism (e.g., average cell volume or leaf surface area), \(\Omega\)
does not have a clear population-level interpretation. In such cases,
eqn 1 still defines eco-evolutionary change; the interpretation of
\(\Omega\) by itself is just not as interesting, biologically.

Box 3 provides an instructive example of three plants with different
fitnesses and carbon captures. We use this example to demonstrate a
concrete example of our proposed framework.

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\begin{quote}
\textbf{Box 3:} As an instructive example of our framework, consider a
population of \(N_{t} = 3\) annual plants in which individual total
fruit mass (kg) is measured, and change is observed over a year. For all
plants, \(\delta_{i} = 1\) (because annual plants die between \(t\) and
\(t+1\)), and let plant fecundities be \(\beta_{1} = 1\),
\(\beta_{2} = 1\), and \(\beta_{3} = 2\). Applying eqn 1 to population
change such that \(\Omega = N_{t+1}\), \(z_{i} = 1\) and
\(\Delta z_{i} = 0\),
\(N_{t+1} = (1 - 1 + 1)(1 + 0) + (1 - 1 + 1)(1 + 0) + (2 - 1 + 1)(1 + 0) = 4\)
(note \(\bar{w} = 4/3\), so \(N_{t}\bar{w} = 3(4/3) = 4\)). Focusing
next on the characteristic of total fruit mass, let \(z_{1} = 0.8\) kg,
\(z_{2} = 1.0\) kg, and \(z_{3} = 1.5\) kg. Also let
\(\Delta z_{i} = 0.1\) for all plants to reflect a change in soil
environment from \(t\) to \(t + 1\). In this case, population fruit
yield at \(t+1\) is
\(\Omega = (1 - 1 + 1)(0.8 + 0.1) + (1 - 1 + 1)(1.0 + 0.1) + (2 - 1 + 1)(1.5 + 0.1) = 5.2\)
kg. At \(t\), mean fruit yield per plant was
\((0.8 + 1.0 + 1.5)/3 = 1.1\) kg, but at \(t+1\), mean fruit yield per
plant is \([(0.8 + 0.1) + (1.0 + 0.1) + 2(1.5 + 0.1)]/4 = 1.3\) kg. Note
that
\(\mathrm{Cov}(w, z) = (1/N)\sum_{i}^{N}(z_{i} - \bar{z})(w_{i} - \bar{w}) = (1/3)\left[(0.8 - 1.1)(1-4/3) + (1 - 1.1)(1-4/3) + (1.5 - 1.1)(2-4/3) \right]  = 2/15\)
and \(\mathrm{E}(w \Delta z) = 4/3 \times 0.1 = 2/15\), so applying the
Price equation, \(\bar{w}\Delta \bar{z} = 2/15 + 2/15 = 4/15\). Since
\(\bar{w} = 4/3\), multiplying both sides of the equation by \(3/4\)
returns \(\Delta z = 0.2\), which is the mean difference in fruit yield
between \(t+1\) and \(t\). The framework expressed in eqn 1 thereby
links population change, evolutionary change, and ecosystem function.
\end{quote}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\section{Discussion}\label{discussion}

An important aspect of scientific progress is the ability to connect
disparate theories and models to show how specific empirical and
theoretical models are logical (mathematical) consequences of more
fundamental ones (\citeproc{ref-Nagel1961}{Nagel 1961};
\citeproc{ref-Morrison2000}{Morrison 2000}). Rather than making
simplifying assumptions, as is the approach for specific ecological and
evolutionary models, we focus on fundamental axioms that are universal
to closed biological systems: discrete individuals, birth, death, and
change over time. We define an abstract sum (\(\Omega\)), to which all
individuals within the population contribute. From the basic axiom that
each individual is one member of a specific population (\(z_{i} = 1\)
and \(\Delta z_{i} = 0\)), we recover the most general equation of
population ecology (Box 2) and the principle of exponential growth
(\(N_{t+1} = N_{t}\lambda\)). By defining individual fitness (\(w_{i}\))
and applying the total conservation of probability to individual
frequencies (\citeproc{ref-Frank2015}{Frank 2015},
\citeproc{ref-Frank2016}{2016}), we recover the most fundamental
equation of evolution, the Price equation (Box 1). Our eqn 1 thereby
provides a foundation for defining eco-evolutionary change in any
population.

The Price equation provides a complete and exact description of
evolution in any closed evolving system (\citeproc{ref-Price1970}{Price
1970}; \citeproc{ref-Frank2012}{Frank 2012}). It is derived by
rearranging the mathematical notation defining changes in the
frequencies and characteristics of any type of entity (e.g.,
individuals, alleles, \citeproc{ref-Price1970}{Price 1970};
\citeproc{ref-Gardner2008}{Gardner 2008}; \citeproc{ref-Luque2016}{Luque
2017}). This derivation partitions total characteristic change into
different components, making it possible to isolate evolutionary
mechanisms (e.g., selection) and levels of biological organisation
(e.g., group, individual) (\citeproc{ref-Frank1995}{Frank 1995},
\citeproc{ref-Frank2012}{2012}; \citeproc{ref-Kerr2009}{Kerr and
Godfrey-Smith 2009}; \citeproc{ref-Luque2016}{Luque 2017};
\citeproc{ref-Okasha2020}{Okasha and Otsuka 2020}). Because of its
abstract nature and lack of any system-specific assumptions, the Price
equation is not dynamically sufficient and makes no predictions about
what will happen in any particular system
(\citeproc{ref-Gardner2020}{Gardner 2020}). Its role is not to predict,
but to formally and completely define and separate components of
evolutionary change (\citeproc{ref-Baravalle2024}{Baravalle et al.
2025}). The same is true of the general equation for population change
(eqn 2), at least as we have used it here where it serves to define what
population growth means in ecology. This equation formally and
completely describes population change in terms of births and deaths. In
eqn 1, we therefore have a fundamental equation from which we can derive
complete ecological and evolutionary change in any closed biological
population. Like all fundamental equations, our equation is necessarily
abstract and not dynamically sufficient. We believe that it will be
useful for eco-evolutionary theory in a similar way that the Price
equation is useful for evolutionary theory: potentially facilitating
specific model development and identifying new conceptual insights,
unresolved errors, and sources of model disagreements (see below and
Supporting Information S1).

Our unification recovers the equivalence between the finite rate of
increase \(\lambda\) (Box 2) and population mean evolutionary fitness
\(\bar{w}\) (Box 1). The population growth equation
\(N_{t+1} = N_{t}\lambda\) can always be rewritten as
\(N_{t+1} = N_{t}\bar{w}\). This specific equivalence has been proposed
before (e.g., \citeproc{ref-Lande1976}{Lande 1976}), as has the broader
relationship between population growth rate and evolutionary fitness
(\citeproc{ref-Fisher1930}{Fisher 1930};
\citeproc{ref-Charlesworth1980}{Charlesworth 1980};
\citeproc{ref-Lande1982}{Lande 1982}; \citeproc{ref-Roff2008}{Roff
2008}; \citeproc{ref-Lion2018}{Lion 2018}). We show this from first
principles and clarify the relationship between fitness and population
growth. Over an arbitrary length of time, fitness is properly defined as
\(w_{i} = \beta_{i} - \delta_{i} + 1\). Over an individual's lifetime
(which, by definition, includes death), fitness is therefore
\(\beta_{i}\). Rates of change in ecology and evolution are reflected in
the first and second statistical moments of fitness, respectively.
Population growth rate reflects mean fitness \(\bar{w}\), while the rate
of evolutionary change reflects the variance in fitness
\(\mathrm{Var}(w)/\bar{w}\) (i.e., Fisher's fundamental theorem,
\citeproc{ref-Frank1997}{Frank 1997}; \citeproc{ref-Rice2004}{Rice
2004}; \citeproc{ref-Queller2017}{Queller 2017}).

Our unification may also help explain, at least partially, some of the
success of classical population genetic models. For decades, population
genetics (and to some extent quantitative genetics) has been accused of
being a reductionist view of evolution, reducing everything to changes
in allele frequencies and abstracting away from individuals and their
environments (the ecological interactions,
\citeproc{ref-MacColl2011}{MacColl 2011}). This has been a line of
argumentation by some defenders of the so-called Extended Synthesis
(\citeproc{ref-Pigliucci2009}{Pigliucci 2009}), especially in relation
to niche construction (\citeproc{ref-Odling-smee2003}{Odling-Smee et al.
2003}). Famously, Mayr (\citeproc{ref-Mayr1960}{1959}) characterised
population genetics as a simple input and output of genes, analogous to
``the adding of certain beans to a beanbag and the withdrawing of
others'' (also called ``beanbag genetics''). Historical critics of
population genetics could not articulate a clear explanation for why it
works so well despite all of its idealisations and simplifications. From
the Price equation, we are able to recover classical population and
quantitative genetic models (\citeproc{ref-Queller2017}{Queller 2017})
and develop new ones (\citeproc{ref-Rice2004}{Rice 2004},
\citeproc{ref-Rice2020}{2020}; \citeproc{ref-Luque2016}{Luque 2017};
\citeproc{ref-Lion2018}{Lion 2018}). Our eqn 1 contains ecology at its
core, and we show how the Price equation logically follows from it after
accounting for absolute population growth (eqn 8). We therefore conclude
that population and quantitative genetic equations contain ecology (no
matter how hidden), and the ecological nature of evolution is implicit
in population and quantitative genetic models.

We have focused on the dynamics of a closed population, and in doing so
leave ecological and evolutionary change attributable to migration for
future work. In population ecology, immigration and emigration can be
incorporated by adding a term for each to the right-hand side of the
equation in Box 2 (\citeproc{ref-Gotelli2001}{Gotelli 2001}). In
evolution, because the Price equation relies on mapping
ancestor-descendant relationships, accounting for migration is more
challenging. Kerr and Godfrey-Smith (\citeproc{ref-Kerr2009}{2009})
demonstrate how the Price equation can be extended to allow for
arbitrary links between ancestors and descendants, thereby extending the
Price equation to allow for immigration and emigration. Frank
(\citeproc{ref-Frank2012}{2012}) presents a simplified version of Kerr
and Godfrey-Smith (\citeproc{ref-Kerr2009}{2009}) that allows some
fraction of descendants to be unconnected to ancestors. In both ecology
and evolution, accounting for migration is done through the use of
additional terms on the right-hand side of the equations.

We believe our fundamental equation to be complete and exact for any
closed population. It therefore implicitly includes any effects of
density dependence on population growth (see Box 2), or any social
effects on evolutionary change (see Box 1). Both of these effects can be
made explicit by specifying how other individuals in a population affect
birth and death of a focal individual. We demonstrate this by deriving
more specific models of density dependent population growth and
multi-level selection in Supporting Information S1. In Supporting
Information S1, we also suggest how new theoretical insights might be
gained by integrating density-dependent growth and social interactions
into our framework.

We have shown that we can derive the fundamental equations of population
ecology and biological evolution from a single unifying equation.
Lastly, we propose our eqn 1 as a potential starting point for defining
ecosystem function and further conceptual unification between ecology,
evolution, and ecosystem function. The Price equation has previously
been used to investigate ecosystem function
(\citeproc{ref-Loreau2001}{Loreau and Hector 2001};
\citeproc{ref-Fox2006}{Fox 2006}), but not with any attempt towards
conceptual unification with evolutionary biology. For example, Fox
(\citeproc{ref-Fox2006}{2006}) applied the abstract properties of the
Price equation to partition total change in ecosystem function into
separate components attributable to species richness, species
composition, and context dependent effects. This approach provides a
framework for comparing the effects of biodiversity on ecosystem
function in empirical systems (\citeproc{ref-Fox2006}{Fox 2006};
\citeproc{ref-Winfree2015}{Winfree et al. 2015};
\citeproc{ref-Mateo-Tomas2017}{Mateo-Tomás et al. 2017}). Instead, our
eqn 1 defines \(\Omega\) as total ecosystem function contributed by a
focal population. It is therefore possible to investigate ecological,
evolutionary, and ecosystem function change from the same shared
framework.

\section{Acknowledgements}\label{acknowledgements}

This manuscript was supported by joint funding between the French
Foundation for Research on Biodiversity (FRB) Centre for the Synthesis
and Analysis of Biodiversity (CESAB) and the German Centre for
Integrative Biodiversity Research (sDiv). It was written as part of the
Unification of Modern Coexistence Theory and Price Equation (UNICOP)
project. Victor J. Luque was also supported by the Spanish Ministry of
Science and Innovation (Project: PID2021-128835NB-I00), and the
Conselleria d'Innovaci\(\'{o}\), Universitats, Ci\(\`{e}\)ncia i
Societat Digital -- Generalitat Valenciana (Project: CIGE/2023/16). We
are grateful for many conversations with S\(\'{e}\)bastien Lion, Kelsey
Lyberger, Swati Patel, and especially Lynn Govaert, whose questions
helped us clarify the relationship between population growth and
fitness. Brad Duthie would also like to thank Matt Tinsley, Brent
Danielson, and Stan Harpole. Victor J. Luque would also like to thank
Lorenzo Baravalle, Pau Carazo, Santiago Ginnobili, Manuel Serra, and
Ariel Roff\(\'{e}\).

\section{Author Contributions}\label{author-contributions}

Both authors came up with the question idea. Following many discussions
between the authors, ABD proposed the initial equation with subsequent
exploration and development from both authors. Both authors contributed
to the writing.

\section{Competing Interests}\label{competing-interests}

The authors declare no competing interests.

\section{Data Availability}\label{data-availability}

This work does not include any data.

\section{References}\label{references}

\phantomsection\label{refs}
\begin{CSLReferences}{1}{0}
\bibitem[\citeproctext]{ref-Baravalle2022}
Baravalle, L., and V. J. Luque. 2022.
\href{https://doi.org/10.1387/theoria.21940}{Towards a pricean
foundation for cultural evolutionary theory}. Theoria 37:209--231.

\bibitem[\citeproctext]{ref-Baravalle2024}
Baravalle, L., A. Roffé, V. J. Luque, and S. Ginnobili. 2025.
\href{https://doi.org/10.1007/s13752-024-00482-4}{The value of price}.
Pages 12--24 \emph{in}Biological Theory (Vol. 20).

\bibitem[\citeproctext]{ref-Barfield2011}
Barfield, M., R. D. Holt, and R. Gomulkiewicz. 2011.
\href{https://doi.org/10.1086/658903}{Evolution in stage-structured
populations}. American Naturalist 177:397--409.

\bibitem[\citeproctext]{ref-Bassar2021}
Bassar, R. D., T. Coulson, J. Travis, and D. N. Reznick. 2021.
\href{https://doi.org/10.1111/ele.13712}{Towards a more precise -- and
accurate -- view of eco-evolution}. Ecology Letters 24:623--625.

\bibitem[\citeproctext]{ref-Brantingham2007}
Brantingham, P. J. 2007. \href{https://doi.org/10.2307/40035853}{{A
unified evolutionary model of archaeological style and function based on
the Price equation}}. American Antiquity 72:395--416.

\bibitem[\citeproctext]{ref-Charlesworth1980}
Charlesworth, B. 1980. Evolution in age-structured populations.
Cambridge studies in mathematical biology. Cambridge University Press,
Cambridge.

\bibitem[\citeproctext]{ref-Collins2009}
Collins, S., and A. Gardner. 2009.
\href{https://doi.org/10.1111/j.1461-0248.2009.01340.x}{{Integrating
physiological, ecological and evolutionary change: A Price equation
approach}}. Ecology Letters 12:744--757.

\bibitem[\citeproctext]{ref-Connor2004}
Connor, J., and D. L. Hartl. 2004. {A premier of ecological genetics}.
Sinauer Associates Incorporated.

\bibitem[\citeproctext]{ref-Coulson2008}
Coulson, T., and S. Tuljapurkar. 2008.
\href{https://doi.org/10.1086/591693}{The dynamics of a quantitative
trait in an age-structured population living in a variable environment}.
American Naturalist 172:599--612.

\bibitem[\citeproctext]{ref-Darwin1859}
Darwin, C. 1859. The origin of species. Penguin.

\bibitem[\citeproctext]{ref-Dobzhansky1970}
Dobzhansky, T. 1970. Genetics of the evolutionary process (Vol. 139).
Columbia University Press.

\bibitem[\citeproctext]{ref-Edelaar2023}
Edelaar, Pim, J. Otsuka, and V. J. Luque. 2023.
\href{https://doi.org/10.1111/brv.12910}{{A generalised approach to the
study and understanding of adaptive evolution}}. Biological Reviews
98:352--375.

\bibitem[\citeproctext]{ref-Ewens2014}
Ewens, W. J. 2014.
\href{https://doi.org/10.1007/s10539-013-9412-0}{Grafen, the price
equations, fitness maximization, optimisation and the fundamental
theorem of natural selection}. Biology and Philosophy 29:197--205.

\bibitem[\citeproctext]{ref-Fisher1930}
Fisher, R. A. 1930. The genetical theory of natural selection. Oxford
University Press, Oxford, UK.

\bibitem[\citeproctext]{ref-Fisher1958}
Fisher, R. A. 1958. The genetical theory of natural selection (2nd ed.).
Dover.

\bibitem[\citeproctext]{ref-Fox2006}
Fox, J. W. 2006.
\href{https://doi.org/10.1890/0012-9658(2006)87\%5B2687:utpetp\%5D2.0.co;2}{{Using
the price equation to partition the effects of biodiversity loss on
ecosystem function}}. Ecology 87:2687--2696.

\bibitem[\citeproctext]{ref-Frank1995}
Frank, S. A. 1995. \href{https://doi.org/10.1006/jtbi.1995.0148}{{George
Price's contributions to evolutionary genetics}}. Journal of Theoretical
Biology 175:373--388.

\bibitem[\citeproctext]{ref-Frank1997}
---------. 1997.
\href{https://doi.org/10.1111/j.1558-5646.1997.tb05096.x}{{The Price
equation, Fisher's fundamental theorem, kin selection, and causal
analysis}}. Evolution 51:1712--1729.

\bibitem[\citeproctext]{ref-Frank2012}
---------. 2012.
\href{https://doi.org/10.1111/j.1420-9101.2012.02498.x}{{Natural
selection. IV. The Price equation}}. Journal of Evolutionary Biology
25:1002--1019.

\bibitem[\citeproctext]{ref-Frank2015}
---------. 2015. \href{https://doi.org/10.3390/e17107087}{{D'Alembert's
direct and inertial forces acting on populations: The price equation and
the fundamental theorem of natural selection}}. Entropy 17:7087--7100.

\bibitem[\citeproctext]{ref-Frank2016}
---------. 2016. \href{https://doi.org/10.3390/e18050192}{{Common
probability patterns arise from simple invariances}}. Entropy 18:1--22.

\bibitem[\citeproctext]{ref-Frank2015a}
---------. 2017. \href{https://doi.org/10.1002/ece3.2922}{{Universal
expressions of population change by the Price equation: natural
selection, information, and maximum entropy production}}. Ecology and
Evolution 1--16.

\bibitem[\citeproctext]{ref-Gardner2008}
Gardner, A. 2008. \href{https://doi.org/10.1016/j.cub.2008.01.005}{{The
Price equation}}. Current Biology 18:198--202.

\bibitem[\citeproctext]{ref-Gardner2020}
---------. 2020. \href{https://doi.org/10.1098/rstb.2019.0361}{{Price's
equation made clear}}. Philosophical Transactions of the Royal Society
B: Biological Sciences 375:20190361.

\bibitem[\citeproctext]{ref-Godsoe2021}
Godsoe, W., K. E. Eisen, D. Stanton, and K. M. Sirianni. 2021.
\href{https://doi.org/10.1007/s12080-020-00478-3}{Selection and
biodiversity change}. Theoretical Ecology 14:367--379.

\bibitem[\citeproctext]{ref-Gotelli2001}
Gotelli, N. J. 2001. A primer of ecology. Sinauer associate. Inc.
Sunderland, MA.

\bibitem[\citeproctext]{ref-Govaert2019}
Govaert, L., E. A. Fronhofer, S. Lion, C. Eizaguirre, D. Bonte, M. Egas,
A. P. Hendry, et al. 2019.
\href{https://doi.org/10.1111/1365-2435.13241}{{Eco-evolutionary
feedbacks---Theoretical models and perspectives}}. Functional Ecology
33:13--30.

\bibitem[\citeproctext]{ref-Jaggi2024}
Jaggi, H., W. Zuo, R. Kentie, J. M. Gaillard, T. Coulson, and S.
Tuljapurkar. 2024. \href{https://doi.org/10.1111/ele.14551}{Density
dependence shapes life-history trade-offs in a food-limited population}.
Ecology letters 27:e14551.

\bibitem[\citeproctext]{ref-Kerr2009}
Kerr, B., and P. Godfrey-Smith. 2009.
\href{https://doi.org/10.1111/j.1558-5646.2008.00570.x}{{Generalization
of the price equation for evolutionary change}}. Evolution 63:531--536.

\bibitem[\citeproctext]{ref-Kitcher1993}
Kitcher, P. 1993. {The advancement of science}. Oxford University Press,
New York.

\bibitem[\citeproctext]{ref-Lande1976}
Lande, R. 1976. Natural selection and random genetic drift in phenotypic
evolution. Evolution 30:314--334.

\bibitem[\citeproctext]{ref-Lande1982}
---------. 1982. \href{https://doi.org/10.2307/1936778}{{A quantitative
genetic theory of life history evolution}}. Ecology 63:607--615.

\bibitem[\citeproctext]{ref-Lehtonen2016}
Lehtonen, J. 2016.
\href{https://doi.org/10.1016/j.tree.2016.07.006}{{Multilevel selection
in kin selection language}}. Trends in Ecology and Evolution xx:1--11.

\bibitem[\citeproctext]{ref-Lehtonen2018}
---------. 2018. \href{https://doi.org/10.1086/694891}{{The Price
equation, gradient dynamics, and continuous trait game theory}}.
American Naturalist 191:146--153.

\bibitem[\citeproctext]{ref-Lehtonen2020a}
---------. 2020. \href{https://doi.org/10.1098/rstb.2019.0362}{The price
equation and the unity of social evolution theory}. Philosophical
Transactions of the Royal Society B: Biological Sciences 375:20190362.

\bibitem[\citeproctext]{ref-Lehtonen2020}
Lehtonen, J., S. Okasha, and H. Helanterä. 2020.
\href{https://doi.org/10.1098/rstb.2019.0350}{{Fifty years of the Price
equation}}. Philosophical Transactions of the Royal Society B:
Biological Sciences 375:20190350.

\bibitem[\citeproctext]{ref-Levins1966}
Levins, R. 1966. \href{https://doi.org/10.2307/27836590}{The strategy of
model building in population biology}. American Naturalist.

\bibitem[\citeproctext]{ref-Lion2018}
Lion, S. 2018. \href{https://doi.org/10.1086/694865}{{Theoretical
approaches in evolutionary ecology: environmental feedback as a unifying
perspective}}. American Naturalist 191.

\bibitem[\citeproctext]{ref-Lion2023}
Lion, S., A. Sasaki, and M. Boots. 2023.
\href{https://doi.org/10.1111/ele.14183}{Extending eco-evolutionary
theory with oligomorphic dynamics}. Ecology Letters 26:S22--S46.

\bibitem[\citeproctext]{ref-Loreau2001}
Loreau, M., and A. Hector. 2001.
\href{https://doi.org/10.1038/35083573}{{Partitioning selection and
complementarity in biodiversity experiments}}. Nature 412:72--76.

\bibitem[\citeproctext]{ref-Luque2016}
Luque, V. J. 2017. \href{https://doi.org/10.1007/s10539-016-9538-y}{{One
equation to rule them all: a philosophical analysis of the Price
equation}}. Biology and Philosophy 32:1--29.

\bibitem[\citeproctext]{ref-Luque2021}
Luque, V. J., and L. Baravalle. 2021.
\href{https://doi.org/10.1007/s11229-021-03339-6}{{The mirror of
physics: on how the price equation can unify evolutionary biology}}.
Synthese 199:12439--12462.

\bibitem[\citeproctext]{ref-MacCallum2012}
MacCallum, R. M., M. Mauch, A. Burt, and A. M. Leroi. 2012.
\href{https://doi.org/10.5061/dryad.h0228}{{Evolution of music by public
choice}}. Proceedings of the National Academy of Sciences
109:12081--12086.

\bibitem[\citeproctext]{ref-MacColl2011}
MacColl, A. D. C. 2011.
\href{https://doi.org/10.1016/j.tree.2011.06.009}{{The ecological causes
of evolution}}. Trends in Ecology and Evolution 26:514--522.

\bibitem[\citeproctext]{ref-Mateo-Tomas2017}
Mateo-Tomás, P., P. P. Olea, M. Moleón, N. Selva, and J. A.
Sánchez-Zapata. 2017. \href{https://doi.org/10.1111/geb.12673}{{Both
rare and common species support ecosystem services in scavenger
communities}}. Global Ecology and Biogeography 26:1459--1470.

\bibitem[\citeproctext]{ref-Mayr1960}
Mayr, E. 1959. \href{https://doi.org/10.1101/SQB.1959.024.01.003}{Where
are we? Genetics and twentieth century darwinism}. Pages 1--14
\emph{in}Cold spring harbor symposia on quantitative biology (Vol. 24).

\bibitem[\citeproctext]{ref-Morrison2000}
Morrison, M. 2000. Unifying scientific theories: Physical concepts and
mathematical structures. Cambridge University Press.

\bibitem[\citeproctext]{ref-Nagel1961}
Nagel, E. 1961. The structure of science: Problems in the logic of
scientific explanation. Harcourt, Brace \& World, New York, NY, USA.

\bibitem[\citeproctext]{ref-NewellEtAl2019}
Newell, D. B., and E. Tiesinga. 2019.
\href{https://doi.org/10.6028/NIST.SP.330-2019}{The international system
of units (SI)}.

\bibitem[\citeproctext]{ref-Odling-smee2003}
Odling-Smee, F. J., K. N. Laland, and M. W. Feldman. 2003. Niche
construction: The neglected process in evolution. Princeton University
Press.

\bibitem[\citeproctext]{ref-Okasha2006}
Okasha, S. 2006.
\href{https://doi.org/10.1093/acprof:oso/9780199267972.001.0001}{Evolution
and the levels of selection}. Oxford University Press.

\bibitem[\citeproctext]{ref-Okasha2020}
Okasha, S., and J. Otsuka. 2020.
\href{https://doi.org/10.1098/rstb.2019.0365}{{The Price equation and
the causal analysis of evolutionary change}}. Philosophical Transactions
of the Royal Society B: Biological Sciences 375:20190365.

\bibitem[\citeproctext]{ref-Page2002}
Page, K. M., and M. A. Nowak. 2002.
\href{https://doi.org/10.1006/jtbi.2002.3112}{Unifying evolutionary
dynamics}. Journal of theoretical biology 219:93--98.

\bibitem[\citeproctext]{ref-Patel2018}
Patel, S., M. H. Cortez, and S. J. Schreiber. 2018.
\href{https://doi.org/10.1101/104505}{{Partitioning the effects of
eco-evolutionary feedbacks on community stability}}. American Naturalist
191:1--29.

\bibitem[\citeproctext]{ref-Pelletier2009}
Pelletier, F., D. Garant, and A. P. Hendry. 2009.
\href{https://doi.org/10.1098/rstb.2009.0027}{Eco-evolutionary
dynamics}. Philosophical Transactions of the Royal Society London B
364:1483--1489.

\bibitem[\citeproctext]{ref-Pigliucci2009}
Pigliucci, M. 2009.
\href{https://doi.org/10.1111/j.1749-6632.2009.04578.x}{{An extended
synthesis for evolutionary biology}}. Annals of the New York Academy of
Sciences 1168:218--228.

\bibitem[\citeproctext]{ref-Price1970}
Price, G. R. 1970. \href{https://doi.org/10.1038/227520a0}{{Selection
and covariance}}.

\bibitem[\citeproctext]{ref-Price1972}
---------. 1972.
\href{https://doi.org/10.1111/j.1469-1809.1957.tb01874.x}{{Extension of
covariance selection mathematics}}. Annals of Human Genetics
35:485--490.

\bibitem[\citeproctext]{ref-Queller2017}
Queller, D. C. 2017. \href{https://doi.org/10.1086/690937}{{Fundamental
theorems of evolution}}. American Naturalist 189:000--000.

\bibitem[\citeproctext]{ref-Rees2016}
Rees, M., and S. P. Ellner. 2016.
\href{https://doi.org/10.1111/2041-210X.12487}{Evolving integral
projection models: Evolutionary demography meets eco-evolutionary
dynamics}. Methods in Ecology and Evolution 7:157--170.

\bibitem[\citeproctext]{ref-Rice2004}
Rice, S. H. 2004. {Evolutionary theory: mathematical and conceptual
foundations}. Sinauer Associates Incorporated.

\bibitem[\citeproctext]{ref-Rice2020}
---------. 2020.
\href{https://doi.org/10.1098/rstb.2019.0353}{{Universal rules for the
interaction of selection and transmission in evolution}}. Philosophical
Transactions of the Royal Society B: Biological Sciences 375.

\bibitem[\citeproctext]{ref-Rice2009}
Rice, S. H., and A. Papadopoulos. 2009.
\href{https://doi.org/10.1371/journal.pone.0007130}{{Evolution with
stochastic fitness and stochastic migration}}. PLoS One 4.

\bibitem[\citeproctext]{ref-Robertson1966}
Robertson, A. 1966. \href{https://doi.org/10.1017/S0003356100037752}{{A
mathematical model of the culling process in dairy cattle}}. Animal
Science 8:95--108.

\bibitem[\citeproctext]{ref-Roff2008}
Roff, D. A. 2008.
\href{https://doi.org/10.1007/s12041-008-0056-9}{{Defining fitness in
evolutionary models}}. Journal of Genetics 87:339--348.

\bibitem[\citeproctext]{ref-Simmonds2020}
Simmonds, E. G., E. F. Cole, B. C. Sheldon, and T. Coulson. 2020.
\href{https://doi.org/10.1111/ele.13603}{Phenological asynchrony: A
ticking time-bomb for seemingly stable populations?} Ecology Letters.
Blackwell Publishing Ltd.

\bibitem[\citeproctext]{ref-Smocovitis1992}
Smocovitis, V. B. 1992. \emph{Unifying biology: The evolutionary
synthesis and evolutionary biology}. Source: Journal of the History of
Biology (Vol. 25).

\bibitem[\citeproctext]{ref-Turchin2001}
Turchin, P. 2001.
\href{https://doi.org/10.1034/j.1600-0706.2001.11310.x}{Does population
ecology have general laws?} Oikos 94:17--26.

\bibitem[\citeproctext]{ref-Ulrich2024}
Ulrich, W., N. J. Gotelli, G. Strona, and W. Godsoe. 2024.
\href{https://doi.org/10.1016/j.ecolmodel.2024.110695}{Reconsidering the
price equation: Benchmarking the analytical power of additive
partitioning in ecology}. Ecological Modelling 491:110695.

\bibitem[\citeproctext]{ref-VanVeelen2005}
van Veelen, M. 2005.
\href{https://doi.org/10.1016/j.jtbi.2005.04.026}{{On the use of the
Price equation}}. Journal of Theoretical Biology 237:412--426.

\bibitem[\citeproctext]{ref-VanVeelen2020}
---------. 2020. \href{https://doi.org/10.1098/rstb.2019.0355}{{The
problem with the Price equation}}. Philosophical Transactions of the
Royal Society B: Biological Sciences 375:20190355.

\bibitem[\citeproctext]{ref-VanVeelen2012}
van Veelen, M., J. García, M. W. Sabelis, and M. Egas. 2012.
\href{https://doi.org/10.1016/j.jtbi.2011.07.025}{{Group selection and
inclusive fitness are not equivalent; the Price equation vs. models and
statistics}}. Journal of Theoretical Biology 299:64--80.

\bibitem[\citeproctext]{ref-Winfree2015}
Winfree, R., J. W. Fox, N. M. Williams, J. R. Reilly, and D. P.
Cariveau. 2015. \href{https://doi.org/10.1111/ele.12424}{{Abundance of
common species, not species richness, drives delivery of a real-world
ecosystem service}}. Ecology Letters 18:626--635.

\bibitem[\citeproctext]{ref-Yamamichi2023}
Yamamichi, M., S. P. Ellner, and N. G. Hairston. 2023.
\href{https://doi.org/10.1111/ele.14197}{Beyond simple adaptation:
Incorporating other evolutionary processes and concepts into
eco-evolutionary dynamics}. Ecology Letters 26:S16--S21.

\end{CSLReferences}

\clearpage

\section{Supporting Information S1}\label{supporting-information-s1}

This supporting information demonstrates how to derive well-established
models in population ecology and evolutionary biology from equation 1 in
the main text,

\[\Omega = \sum_{i=1}^{N} \left(\beta_{i} - \delta_{i} + 1 \right)\left(z_{i} + \Delta z_{i} \right).
\tag{1}
\]

In the main text, we derived both the Price equation and the birth-death
model from the above. Here we integrate interactions between individuals
to recover density-dependent population growth, and we integrate groups
within the population to recover multi-level selection. Finally, we
integrate both to model a system in which multi-level selection and
density dependent population change occur simultaneously. We can do this
in a way that clarifies model assumptions by allowing an individual
\(j\) to modulate the birth or death of the focal individual \(i\).

\section{Density-dependent population
growth}\label{density-dependent-population-growth}

There are two potential ways to model the incorporation of density
dependence into population growth. We start with what is likely the most
familiar model focusing on individual growth rate \(r_{i}\), then use a
slightly different model focusing on fitness \(w_{i}\). First note that
here we set \(\Omega = N_{t+1}\), and \(z_{i} = 1\) and
\(\Delta z_{i} = 0\) for all individuals as in the main text. We can
define \(r_{i} = \beta_{i} - \delta_{i}\) as an individual growth rate
for \(i\) (\citeproc{ref-Lion2018}{Lion 2018};
\citeproc{ref-Lion2023}{Lion et al. 2023}). In this case,

\[N_{t+1} = \sum_{i=1}^{N_{t}}\left(r_{i} + 1\right)
\tag{S1}
\]

Mathematically, the most general approach here would be to define
individual growth as a function of the entire system \(\mathbf{E}\),
\(r_{i}(\mathbf{E})\), where \(\mathbf{E}\) is a vector with elements
including any parameters potentially relevant to \(r_{i}\). Taking this
approach would recover a version of eqn 2 in Lion
(\citeproc{ref-Lion2018}{2018}) and permit any relationship between the
system and a focal individual's growth. Limiting our focus to the
effects of other individuals (\(j\)) and assuming that the effects of
these individuals are additive, let \(a_{ij|\cdot}\) be the effect of
individual \(j\) on the growth rate attributable to \(i\) conditioned on
all other individuals within the population such that
\(r_{i}\left(1 + \sum_{i = j}^{N}a_{ij|\cdot} \right)\) defines the
realised growth rate of \(i\),

\[N_{t+1} = \sum_{i=1}^{N_{t}}\left(r_{i}\left(1 - \sum_{i = j}^{N_{t}}a_{ij|\cdot} \right) + 1\right).
\tag{S2}
\]

Assuming that individual effects of \(j\) on \(i\) are also independent,
we can remove the condition,

\[N_{t+1} = \sum_{i=1}^{N_{t}}\left(r_{i}\left(1 - \sum_{i = j}^{N_{t}}a_{ij} \right) + 1\right).\]

Further assuming that all individuals have the same per capita effect
such that \(a = a_{ij}\) for any \(i\) and \(j\) pair (as might be
reasonable given resource competition in a well-mixed population),

\[N_{t+1} = \sum_{i=1}^{N_{t}}\left(r_{i}\left(1 - a N_{t} \right) + 1\right).\]

If \(r_{i}\) values are identical,

\[N_{t+1} = N_{t} + r N_{t}\left(1 - a N_{t} \right).
\tag{S3}
\]

Equation S3 therefore recovers a classic version of a discrete time
logistic growth by making assumptions from an exact model of
eco-evolutionary change.

An alternative approach would be to define model the effects of an
individual \(j\) on the fitness of \(i\) (\(w_{i}\)), thereby replacing
eqn S1 with \(N_{t+1} = \sum_{i=1}^{N_{t}}w_{i}\) and replacing eqn S2
with,

\[N_{t+1} = \sum_{i=1}^{N_{t}}w_{i}\left(1 - \sum_{i = j}^{N_{t}}\alpha_{ij|\cdot} \right).
\]

Note that we have used \(\alpha_{ij}\) to represent the effect of \(j\)
on the fitness of \(i\) for clarity in the sections below. By making the
same assumptions of additivity, independence, and identical effects such
that \(\alpha = \alpha_{ij}\) for all \(j\) on \(i\), and assuming
fitness is equal (\(w_{i} = \bar{w}\)), we can derive,

\[N_{t+1} = \bar{w}N_{t}(1 - \alpha N_{t}).
\tag{S4}
\]

This is an alternative way to express logistic growth.

\section{Multi-level selection}\label{multi-level-selection}

We can recover multi-level selection from our eqn 1. Here we derive the
original form of the multi-level Price (\citeproc{ref-Price1972}{1972})
equation as it appears in eqn 3.1 of Lehtonen
(\citeproc{ref-Lehtonen2020a}{2020}). Individuals belong to one of \(K\)
total groups where \(j\) indexes \(K\) groups and \(i\) indexes
individuals. Individuals do not overlap in group membership. The size of
group \(j\) is denoted as \(N_{j}\). Equation S5 below uses summations
to partition how individuals within each group contribute to \(\Omega\),

\[\Omega = \sum_{j=1}^{K}\sum_{i=1}^{N_{j}}\left(\beta_{j,i} - \delta_{j,i} + 1 \right)\left(z_{j, i} + \Delta z_{j, i}\right).
\tag{S5}
\]

In S5, indices \(\beta_{j, i}\), \(\delta_{j, i}\), and \(z_{j, i}\)
identify individual \(i\) in group \(j\). We set
\(w_{j, i} = \beta_{j, i} - \delta_{j, i} + 1\), and for simplicity let
\(\Delta z_{j, i} = 0\) (i.e., no transmission bias),

\[\Omega = \sum_{j=1}^{K}\sum_{i=1}^{N_{j}}w_{j,i}z_{j, i}.\]

For ease of presentation, with no loss of generality, we assume all
group sizes are equal with a group size of \(N_{j} = n\) for all \(j\).
If group sizes differ, then weighted expectations and covariances are
instead needed (\citeproc{ref-Lehtonen2020a}{Lehtonen 2020}). Given
equal group sizes, the total number of individuals (\(N\)) equals
\(K \times n\), and,

\[\frac{\Omega}{K n} = \left(\frac{1}{K}\right)\left(\frac{1}{n}\right)\sum_{j=1}^{K}\sum_{i=1}^{n}w_{j,i}z_{j, i}.\]

Rearranging,

\[\frac{\Omega}{K n} = \frac{1}{K}\sum_{j=1}^{K}\frac{1}{n}\sum_{i=1}^{n}w_{j,i}z_{j, i}.\]

The inner summation can be rewritten as an expectation for group \(j\),

\[\frac{\Omega}{K n} = \frac{1}{K}\sum_{j=1}^{K}\mathrm{E}_{j}\left(w_{j,i} z_{j,i}\right).\]

As in the main text, we note
\(\mathrm{E}(XY) = \mathrm{Cov}(X,Y) + \mathrm{E}(X)\mathrm{E}(Y)\).
Defining \(Cov_{j}(w_{j}, z_{j})\) as the covariance between \(w_{j,i}\)
and \(z_{j, i}\) for group \(j\),

\[\frac{\Omega}{K n} = \frac{1}{K}\sum_{j=1}^{K}\mathrm{Cov}_{j}\left(w_{j,i}, z_{j,i}\right) + \mathrm{E}_{j}\left(w_{j,i}\right)\mathrm{E}_{j}\left(z_{j,i}\right).\]

We can separate the summation for each term,

\[\frac{\Omega}{K n} = \frac{1}{K}\sum_{j=1}^{K}\mathrm{Cov}_{j}\left(w_{j,i}, z_{j,i}\right) + \frac{1}{K}\sum_{j=1}^{K}\mathrm{E}_{j}\left(w_{j,i}\right)\mathrm{E}_{j}\left(z_{j,i}\right).\]

Using the notation \(\bar{w}_{j} = E_{j}(w_{j,i})\) and
\(\bar{z}_{j} = E_{j}(z_{j,i})\) to indicate the expectation in group
\(j\),

\[\frac{\Omega}{K n} = \frac{1}{K}\sum_{j=1}^{K}\mathrm{Cov}_{j}\left(w_{j,i}, z_{j,i}\right) + \mathrm{E}\left(\bar{w}_{j} \bar{z}_{j} \right).\]

We can also rewrite the first term on the right-hand side as an
expectation,

\[\frac{\Omega}{K n} = \mathrm{E}\left(\mathrm{Cov}_{j}\left(w_{j}, z_{j}\right)\right) + \mathrm{E}\left(\bar{w}_{j} \bar{z}_{j} \right).\]

We can rearrange the second term on the right-hand side
(\(\bar{\bar{w}}\) indicates grand mean over all groups),

\[\frac{\Omega}{K n} = \mathrm{E}\left(\mathrm{Cov}_{j}\left(w_{j}, z_{j}\right)\right) +  \mathrm{Cov}\left(\bar{w}_{j}, \bar{z}_{j} \right) + \bar{\bar{w}}\bar{\bar{z}}.\]

As in the main text, note that \(Kn\bar{\bar{w}}\) accounts for
differences in total population size from \(t\) to \(t+1\), with
\(\bar{\bar{w}}\) being mean fitness across all groups. We can therefore
set \(\Omega = Kn\bar{\bar{w}}\bar{\bar{z'}}\), so,

\[\frac{Kn\bar{\bar{w}}\bar{\bar{z'}}}{K n} - \bar{\bar{w}}\bar{\bar{z}} = \mathrm{E}\left(\mathrm{Cov}_{j}\left(w_{j}, z_{j}\right)\right) +  \mathrm{Cov}\left(\bar{w}_{j}, \bar{z}_{j} \right).\]

Because \(\Delta \bar{\bar{z}} = \bar{\bar{z'}} - \bar{\bar{z}}\),

\[\bar{\bar{w}}\Delta \bar{\bar{z}} = \mathrm{Cov}\left(\bar{w}_{j}, \bar{z}_{j} \right) + \mathrm{E}\left(\mathrm{Cov}_{j}\left(w_{j}, z_{j}\right)\right).
\tag{S6}
\]

This recovers the multi-level Price (\citeproc{ref-Price1972}{1972})
equation (\citeproc{ref-Lehtonen2020a}{Lehtonen 2020}) from a starting
point of eco-evolutionary change in different groups. Equation S6 can be
found in Lehtonen (\citeproc{ref-Lehtonen2016}{2016}) B2.I, who then
derives a multi-level selection version of Hamilton's rule predicting
the evolution of altruism.

\section{Integration of ecology and
evolution}\label{integration-of-ecology-and-evolution}

For simplicity, we now focus on showing an integration between ecology
and evolution using a population with no multi-level selection and let
\(\Delta z_{j, i} = 0\) (i.e., no transmission bias). As above in the
section on density-dependent population growth, we define
\(w_{i} = \beta_{i} - \delta_{i} + 1\) and use \(\alpha_{i,j}\) to
represent the effect of \(j\) on the fitness of \(i\). Our starting
equation is therefore,

\[\Omega = \sum_{i=1}^{N}w_{i}\left(1 - \sum_{j=1}^{N}\alpha_{i, j}\right)z_{i}.
\tag{S7}
\]

We have already demonstrated that if we assume all individuals have the
same effect on a focal individual such that \(\alpha = \alpha_{i,j}\)
for all \(i\) and \(j\) pairs, we can recover equation S4 when
\(z_{i} = 1\) and \(\Omega\) is therefore interpreted as the count of
entities,

\[N_{t+1} = \bar{w} N_{t} \left(1 - \alpha N_{t}\right).\]

We now start from S7 to derive \(\Delta \bar{z}\). The objective is to
use our definition of eco-evolutionary change to simultaneously recover
how interactions between individuals affect population change and
evolutionary change.

We start by dividing both sides of S7 by \(N\),

\[\frac{\Omega}{N} = \frac{1}{N}\sum_{i=1}^{N}w_{i}\left(1 - \sum_{j=1}^{N}\alpha_{i, j}\right)z_{i}.
\tag{S8}
\]

We can express the right-hand side of eqn S8 as an expectation,

\[\frac{\Omega}{N} = \mathrm{E}\left(w_{i}\left(1 - \sum_{j=1}^{N}\alpha_{i, j}\right)z_{i}\right).
\tag{S9}
\]

We can rewrite the right-hand side in terms of covariances,

\[\frac{\Omega}{N} = \mathrm{Cov}\left(w_{i}\left(1 - \sum_{j=1}^{N}\alpha_{i, j}\right), z_{i}  \right) + \mathrm{E}\left(w_{i}\left(1 - \sum_{j=1}^{N}\alpha_{i, j}\right)\right)\mathrm{E}\left(z_{i}\right). 
\tag{S9}
\]

The expectations in the second term on the right-hand side can be
replaced with overbars to represent the mean,

\[\frac{\Omega}{N} = \mathrm{Cov}\left(w_{i}\left(1 - \sum_{j=1}^{N}\alpha_{i, j}\right), z_{i}  \right) + \overline{w_{i}\left(1 - \sum_{j=1}^{N}\alpha_{i, j}\right)}    \bar{z}_{i}. 
\tag{S9}
\]

In the main text, we noted that \(\Omega = N\bar{w}\bar{z}'\). Here mean
fitness incorporates individual interactions, therefore,

\[\Omega = N\overline{w_{i}\left(1 - \sum_{j=1}^{N}\alpha_{i, j}\right)}\bar{z}'.
\tag{S10}
\]

We can therefore rewrite S9,

\[\overline{w_{i}\left(1 - \sum_{j=1}^{N}\alpha_{i, j}\right)}\bar{z}' -  \overline{w_{i}\left(1 - \sum_{j=1}^{N}\alpha_{i, j}\right)}    \bar{z}_{i} = \mathrm{Cov}\left(w_{i}\left(1 - \sum_{j=1}^{N}\alpha_{i, j}\right), z_{i}  \right). 
\tag{S11}
\]

Noting again \(\Delta \bar{z} = \bar{z}' - \bar{z}\),

\[\overline{w_{i}\left(1 - \sum_{j=1}^{N}\alpha_{i, j}\right)}\Delta\bar{z} = \mathrm{Cov}\left(w_{i}\left(1 - \sum_{j=1}^{N}\alpha_{i, j}\right), z_{i}  \right). 
\tag{S12}
\]

We can rewrite the left-hand side of S12,

\[\overline{w_{i}\left(1 - \sum_{j=1}^{N}\alpha_{i, j}\right)}\Delta\bar{z} = \mathrm{Cov}\left(w_{i} - w_{i}\sum_{j=1}^{N}\alpha_{i, j}, z_{i}  \right). 
\tag{S13}
\]

The covariance term can be split without any additional assumptions,

\[\overline{w_{i}\left(1 - \sum_{j=1}^{N}\alpha_{i, j}\right)}\Delta\bar{z} = \mathrm{Cov}\left(w_{i}, z_{i}  \right) - \mathrm{Cov}\left(w_{i}\sum_{j=1}^{N}\alpha_{i, j}, z_{i}  \right). 
\tag{S14}
\]

If we are able to further assume that \(w_{i}\) and the summation over
\(\alpha_{i,j}\) are independent, then we could rewrite,

\[\overline{w_{i}\left(1 - \sum_{j=1}^{N}\alpha_{i, j}\right)}\Delta\bar{z} = \mathrm{Cov}\left(w_{i}, z_{i}  \right) - \mathrm{Cov}\left(\sum_{j=1}^{N}\alpha_{i, j}, z_{i}  \right)\bar{w_{i}}. 
\tag{S15}
\]

Partitioning fitness into different components with the Price equation
is commonplace. But this derivation highlights, e.g., the ecological and
evolutionary relationship between nonsocial and social components of
fitness, and population size. For example, the second term on the
right-hand side of S13 shows the covariance between the sum of social
interactions and a trait. When traits covary with the interaction
between sociality and fitness, they will have a stronger effect on trait
change. The magnitude of this second term will also increase with \(N\),
which reflects the stronger effect of the interaction between sociality
and fitness when there are more individuals interacting with a focal
individual.

\section{Supporting Information
References}\label{supporting-information-references}

\addcontentsline{toc}{section}{References}

\phantomsection\label{refs}

\begin{CSLReferences}{1}{0}
\bibitem[\citeproctext]{ref-Lehtonen2016}
Lehtonen, Jussi. 2016. {``{Multilevel selection in kin selection
language}.''} \emph{Trends in Ecology and Evolution} xx: 1--11.
\url{https://doi.org/10.1016/j.tree.2016.07.006}.

\bibitem[\citeproctext]{ref-Lehtonen2020a}
---------. 2020. {``The Price Equation and the Unity of Social Evolution
Theory.''} \emph{Philosophical Transactions of the Royal Society B:
Biological Sciences} 375: 20190362.
\url{https://doi.org/10.1098/rstb.2019.0362}.

\bibitem[\citeproctext]{ref-Lion2018}
Lion, Sébastien. 2018. {``{Theoretical approaches in evolutionary
ecology: environmental feedback as a unifying perspective}.''}
\emph{American Naturalist} 191 (1).
\url{https://doi.org/10.1086/694865}.

\bibitem[\citeproctext]{ref-Lion2023}
Lion, Sébastien, Akira Sasaki, and Mike Boots. 2023. {``Extending
Eco-Evolutionary Theory with Oligomorphic Dynamics.''} \emph{Ecology
Letters} 26 (September): S22--46.
\url{https://doi.org/10.1111/ele.14183}.

\bibitem[\citeproctext]{ref-Price1972}
Price, George R. 1972. {``{Extension of covariance selection
mathematics}.''} \emph{Annals of Human Genetics} 35 (4): 485--90.
\url{https://doi.org/10.1111/j.1469-1809.1957.tb01874.x}.

\end{CSLReferences}

\end{document}