### 2025-06-18 foundations_of_community_ecology.Rmd
Basic structure to later be written out

---
title: "Foundations of community ecology: Derivation of community Price equation"
author: |
  | UNICOP Group
date: '18 JUN 2025'
output:
  pdf_document:
      keep_tex: yes
  html_document:
    df_print: paged
  word_document: default
header-includes:
- \usepackage{fancyhdr}
bibliography: refs.bib
link-citations: yes
linkcolor: blue
biblio-style: apalike
---

References 
==========================